### 2020-02-20 TaskBoard.Rmd
Prepare the repo with required structure for checkin 3

---
title: "TaskBoard"
author: "A group of fabulous contributors"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This will be the starting point of generating the required files for submission. Feel free to make your edits here. It serves as a helper to handle the housekeeping of this repository and work distribution. There will be another R Markdown file (paper.rmd) aiming at generating the main delivery, target figures.

## Deliverables of Checkin 3
https://canvas.uw.edu/courses/1354201/assignments/5218657

- a folder called __analysis__ that contains an R Markdown document called 'paper.Rmd' that can be knit to produce the target figures that you identified as the goal of your replication. This 'analysis' folder may also have R script files that are used by your R Markdown file.  It should not have any files that are not used by the R Markdown file. 
- a folder called __data__ that contains your data files, or a README.md that tells the reader where the data can be obtained. 
- If necessary, a folder called __notes__ containing scraps of code and other bits and pieces that you've generated along the way, but are not strictly required for your replication (i.e. not used by the R Markdown file).  We should not have to look in here or run any code in here to reproduce the results in your Rmd. 
- Your main README.md must have two sections near the bottom with heading "Contributing" that has the text "We welcome contributions from everyone. Before you get started, please see our contributor guidelines. Please note that this project is released with a Contributor Code of Conduct. By participating in this project you agree to abide by its terms."  Add these files by running `usethis::use_code_of_conduct()` and `usethis::use_tidy_contributing()` at the R console on your computer. Once you have made these files, edit them files to suit your situation, and add links to the appropriate words in text under the "Contributing"  heading in your READMEmd. to direct readers to those new files on your repo. 
- A commit history that shows an even distribution of contributions from all team members, with relevant and non-cryptic commit messages.

```{r}
usethis::use_code_of_conduct()
```

```{r}
usethis::use_tidy_contributing()
```

### 2020-02-20 analysis/paper.Rmd
Prepare the repo with required structure for checkin 3

---
title: "paper"
author: "A group of fabulous contributors"
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This is a R Markdown file which can be knit to produce the target figures as the goal of our replication. This __analysis__ folder also have R script files that are used by the R Markdown file.  It should not have any files that are not used by the R Markdown file.

## Goal

## Specification

## Summary

## Open Questions

### 2020-02-20 Proposal.Rmd
Add GoogleDoc link to TaskDefinition

--- 2020-02-20 TaskBoard.Rmd
@@
 ---
-title: "TaskBoard"
+title: "TaskDefinition"
 author: "A group of fabulous contributors"
@@
 This will be the starting point of generating the required files for submission. 
-Feel free to make your edits here.
-It serves as a helper to handle the housekeeping of this repository and work distribution.
-There will be another R Markdown file (paper.rmd) aiming at generating the main delivery, target figures.
+Feel free to make your edits here. It serves as a helper to handle the housekeeping of this repository 
+and work distribution. There will be another R Markdown file (paper.rmd) aiming at generating the main delivery, target figures.
+
+Given we will be collaborating on a single Rmd, I created a separate Google Doc at [here](https://docs.google.com/document/d/1O7t3q8Aik00dzFaNpWyqoNlxUk6Z0HkJLZuFpJmjpv4/edit?usp=sharing) 
+just in case we would like to do parallel writing.
 

### 2020-02-20 analysis/paper.Rmd
edit paper

@@
 ---
-title: "paper"
-author: "A group of fabulous contributors"
-date: "2/19/2020"
-output: html_document
+Title: "A replication of a simulation study on measuring and correcting bias of under-5 mortality estimation in populations affected by HIV/AIDS"
+Author:
+  - name: Name Surname
+    affiliation: 1
+  - name: Name Surname,
+    affiliation: 2, 3
+Address:
+  - code:    1
+    address: Affiliation Dept/Program/Center, Institution Name, City, State, Country
+  - code:    2
+    address: Affiliation Dept/Program/Center, Institution Name, City, State, Country
+  - code:    3
+    address: Affiliation Dept/Program/Center, Institution Name, City, State, Country
+Contact:
+  - corresponding-author@mail.com
+Editor:
+  - Name Surname
+Reviewer:
+  - Name Surname
+  - Name Surname
+Publication:
+  received:  Sep, 1, 2015
+  accepted:  Sep, 1, 2015
+  published: Sep, 1, 2015
+  volume:    "**1**"
+  issue:     "**1**"
+  date:      Sep 2015
+  number: 1
+Repository:
+  article:   
+  code:      
+  data:      
+  notebook:  
+Reproduction:
+  - "Original article (title, authors, journal, doi)"
+bibliography:
+  bibliography.bib
+output:
+  output: word_document
 ---
+
+# Introduction
+This is a R Markdown file which can be knit to produce the target figures as the goal of our replication.
+This __analysis__ folder also have R script files that are used by the R Markdown file.
+It should not have any files that are not used by the R Markdown file.
 
@@
 
-## Overview
-This is a R Markdown file which can be knit to produce the target figures as the goal of our replication.
-This __analysis__ folder also have R script files that are used by the R Markdown file.
-It should not have any files that are not used by the R Markdown file.
+```{r}
+x <- 1 + 1
+```
 
-## Goal
+## ...
 
-## Specification
+## ...
 
-## Summary
+## Result
 
-## Open Questions
+## Conclusion
+
+## References cited

### 2020-02-21 R/004_regression_revised.R
make fig4 and fig5

@@ -114,8 +114,8 @@ packageVersion("tidyverse")
 
 # Get data for initial population 22,500
 
-load("./results/regdata/p22500/regdata_all.Rdata")
-load("./results/models/inputs.RData")
+load("./results/regdata_all.Rdata")
+load("./results/inputs.RData")
 
 
 
@@ -541,7 +541,7 @@ out_table <-data.frame()
 countries <- c(figure4="Malawi",figure5="Tanzania")
 
 for(fig_name in names(countries)){
-  fig_name <- "figure5"
+  #fig_name <- "figure5"
   fig_country <- countries[fig_name]
   
   # subset each country
@@ -628,7 +628,7 @@ for(fig_name in names(countries)){
     set_colnames("PredictedAdj")
   
   
-  # Bootstrap confidente intervals for predictions
+  # Bootstrap confidence intervals for predictions
   if(is.null(boot_outs[[fig_name]])){
     # Get output vector for glmnet
     
@@ -653,7 +653,7 @@ for(fig_name in names(countries)){
     y_star <- sapply(seq_len(length(tt)), function(i) ifelse(runif(n) <= pi_hat, 1, 0))
     suff_stat <- t(y_star) %*% x
     
-    bcapar(t0=t0,tt=tt,bb=suff_stat)
+    #bcapar(t0=t0,tt=tt,bb=suff_stat)
     
     # This is the function that we will use for each bootstrap sample
     theta <- function(xy,nx,glmnet_alpha,lambda){
@@ -682,7 +682,7 @@ for(fig_name in names(countries)){
       
       set.seed(600)
       
-      bcajack(x=Xy,B=1000,func=theta,new_data,1.0,lambda)
+      bcajack(x=Xy,B=500,func=theta,new_data,1.0,lambda)
     })
   }
   

### 2020-02-21 analysis/paper.Rmd

--- 2020-02-20 analysis/paper.Rmd
@@
 ---
-title: "paper"
-author: "A group of fabulous contributors"
-date: "2/19/2020"
-output: html_document
+title: "A replication of a simulation study on measuring and correcting bias of under-5 mortality estimation in populations affected by HIV/AIDS"
+author: "Sihao Miao, Ruian Yang, Xiaolu Qian, Jiyu Wang, David Wei"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+Repository:
+  article:   
+  code:      
+  data:      
+  notebook:  
+reproduction:
+  - "Quattrochi, John, et al.
+“Measuring and Correcting Bias in Indirect Estimates of under-5 Mortality in Populations Affected by HIV/AIDS: a Simulation Study.” BMC Public Health, vol.
+19, no.
+1, Dec.
+2019, doi:10.1186/s12889-019-7780-3."
+bibliography:
+  bibliography.bib
+output:
+  output: bookdown::word_document2
 ---
+
+# Introduction
+This is a R Markdown file which can be knit to produce the target figures as the goal of our replication.
+This __analysis__ folder also have R script files that are used by the R Markdown file.
+It should not have any files that are not used by the R Markdown file.
 

### 2020-02-21 analysis/paper.Rmd
add introduction to paper

@@
 # Introduction
-This is a R Markdown file which can be knit to produce the target figures as the goal of our replication.
-This __analysis__ folder also have R script files that are used by the R Markdown file.
-It should not have any files that are not used by the R Markdown file.
+Generalized linear model (GLM) is a widely used method in research.
+It is well-established with high flexibility and relatively straightforward interpretation.
+However, GLMs usually require making assumptions and having the underlying assumptions not hold could invalidate the whole model.
+
+The application of GLM can be observed in many fields including public health.
+@quattrochi2019 introduced an GLM via penalized maximum likelihood to correct the under-5 mortality (U5M) in populations affected by HIV/AIDS.
+This method was developed since the U5M is commonly estimated using survey-based approaches in less developed regions and it is based on a fundamental assumption that the mother's survival is not related to her children's, which is often violated in populations affected by HIV/AIDS.
+Therefore, the author sought a method to correct this bias using GLM.
+
+In our work, we present our endeavor to replicate the model building and regression analysis introduced in @quattrochi2019.
+In the meantime, we attempted using different parameters in both constructing the model and applying it to test how robust the original method was.
+We will in this paper report our method and results from the replication attempt.
 

### 2020-02-21 analysis/paper.Rmd
add Overview section to paper.Rmd and edit bib file

@@
 ```
+# Methods
 
-```{r}
-x <- 1 + 1
-```
+## Overview
+The original model was implemented in R (v3.5.2) [team2013r] with web application (https://johnquattrochi.com/bias/) to show that indirect method such as surveying can underestimate U5M by 0-41% in populations where 0-40% of the people are affected by HIV/AIDS.
+The model was written using the R language [team2013r] and some of the main packages used were tidyverse (v1.2.1) and foreach (v1.4.4).
+The original implementation utilized the doReids package (v1.1.1) that offers a back-end server for parallel processing that was not adopted in our replication due to the limitation in resources.
+In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we reduced the number of populations used in the simulation input from 4480 to 1000.
+The code provided the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented.
+However, we discovered some mistakes regarding output inconsistency and wrong variable names.
+Despite the debugging process, we were able to replicate the implementation.
 
-## ...
+## Model Description
 

### 2020-02-23 analysis/paper.Rmd
Add details in Overview section

@@
 The model was written using the R language [team2013r] and some of the main packages used were tidyverse (v1.2.1) and foreach (v1.4.4).
-The original implementation utilized the doReids package (v1.1.1) that offers a back-end server for parallel processing that was not adopted in our replication due to the limitation in resources.
-In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we reduced the number of populations used in the simulation input from 4480 to 1000.
-The code provided the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented.
+The original implementation utilized the doReids package (v1.1.1) that launches a back-end server in Docker for parallel processing that was not adopted in our replication.
+The main limitation is the computing resource we had for this project within a short period of time.
+Based on experiments from multiple team members, finishing the first simulation step would require an 8-core CPU machine for 2 weeks.
+In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we made a decision to reduce the sample size of our model -- the number of populations used in the simulation input has been reduced from 4480 to 1000.
+The code provided by the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented.
 However, we discovered some mistakes regarding output inconsistency and wrong variable names.

### 2020-02-23 analysis/paper.Rmd
add methods to rmd

--- 2020-02-21 analysis/paper.Rmd
@@
 
+```{r, load-packages}
+
+```
+
 ## Overview
@@
 
-## ...
+## Simulation Construction
 
-## Result
+## Model Fitting and Model Selection
 
-## Conclusion
+Newly-generated simulation data was then imported and fitted to various models to select for a best-performing model which provides least errors.
 
-## References cited
+The simulated data was first divided into training set (80%) and test set (20%).
+The training set was fitted to eight different models, which were further evaluated by calculating several error measures (root mean square error, root median square error, mean relative error and median relative error) on both training set (in-sample) and test set (out-of-sample) shown in Table 4.
+
+The glmnet with alpha = 1 was selected as the best-performing model based on its error measures and was used in predictions later.
+
+```{r, model-fitting-and-selection, include=FALSE}
+
+# Get data from simulation
+load("./results/regdata_all.Rdata")
+load("./results/inputs.RData")
+
+# add agegroup variable
+nbd2k %<>% mutate(agegroup=rep(c(1:7),nrow(nbd2k)/7))
+
+# Generate abs error and rel error
+nbd2k %<>% mutate(abs.err=fiveq0_surv-fiveq0_hiv,
+                  rel.err=(fiveq0_surv-fiveq0_hiv)/fiveq0_hiv
+)
+
+# new dep var 
+nbd2k %<>% mutate(corr=fiveq0_hiv-fiveq0_surv,
+                  corr2=corr+2.100105e-05)
+
+##################################################################
+##### MODEL FITTING AND MODEL SELECTION #####
+##################################################################
+
+compute_error_measures <- function(y,prediction,method,sample){
+
+  # Absolute error
+  abs.error <- abs(y-prediction)
+
+  # Relative error
+  relative.error <- abs.error/abs(y)
+
+  # Root mean square error
+
+  rmse <- sqrt(mean(abs.error^2))
+
+  # Root median square error
+
+  rmedse <- sqrt(median(abs.error^2))
+
+  # Mean relative error
+
+  mre <- mean(Filter(function(x) !is.infinite(x),relative.error),na.rm=TRUE)
+
+  # Median relative error
+
+  medre <- median(relative.error)
+
+  data.frame(method=method,sample=sample,rmse=rmse,rmedse=rmedse,mre=mre,medre=medre,stringsAsFactors = FALSE)
+
+}
+
+# Prepare data
+to.model <- nbd2k %>% mutate(
+  agegroup=as.factor(agegroup))
+
+full.model.formula <- corr ~ fiveq0_surv + agegroup + agegroup*hiv1990 + agegroup*hiv2000+  agegroup*hiv2010 + 
+  agegroup*art_prev2005+ agegroup*art_prev2007+ agegroup*art_prev2009 + tfr2000 + tfr2010
+
+minimum.model.formula <- corr ~ fiveq0_surv
+
+# Out of sample data
+fullsamp <- unique(to.model$i)
+set.seed(400)
+
+# dataset with 80% of full sample
+ns1 <- sample(fullsamp, length(fullsamp)*.8, replace=FALSE)
+
+# Select those i's from nbd2k
+train.sample <- to.model %>% filter(i %in% ns1)
+
+# dataset with 20% of full sample
+tw1 <- setdiff(fullsamp,ns1)
+test.sample <- to.model %>% filter(i %in% tw1)
+
+error_table <- data.frame()
+
+##### Full linear model #####
+
+fit <- lm(full.model.formula,data=train.sample)
+prediction <- predict(fit,newx = train.sample)
+error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Full Linear Model","in-sample"))
+
+prediction <- predict(fit,newx = test.sample)
+error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Full Linear Model","out-of-sample"))
+
+##### Forward and Backward selection #####
+
+# maximum model
+
+model.max <- glm(formula=full.model.formula,
+                 family = "gaussian"(link="identity"), data=train.sample)
+
+# minimum model
+
+model.min <- glm(formula=minimum.model.formula, family = "gaussian"(link="identity"), data=train.sample)
+
+# Forward BIC
+
+fit <- step(object=model.min, scope=list(upper=model.max,lower=~1), direction="forward",k=log(nrow(train.sample)))
+
+prediction <- predict(fit,newx = train.sample)
+
+error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Forward Sel. BIC","in-sample"))
+
+prediction <- predict(fit,newx = test.sample)
+
+error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Forward Sel. BIC","out-of-sample"))
+
+# Forward AIC
+
+fit <- step(object=model.min, scope=list(upper=model.max,lower=~1), direction="forward",k=2)
+
+prediction <- predict(fit,newx = train.sample)
+
+error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Forward Sel. AIC","in-sample"))
+
+prediction <- predict(fit,newx = test.sample)
+
+error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Forward Sel. AIC","out-of-sample"))
+
+# Backward BIC
+
+fit <- step(model.max, direction="backward",k=log(nrow(to.model)))
+
+prediction <- predict(fit,newx = train.sample)
+
+error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Backward Sel. BIC","in-sample"))
+
+prediction <- predict(fit,newx = test.sample)
+
+error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Backward Sel. BIC","out-of-sample"))
+
+# Backward AIC
+
+fit <- step(model.max, direction="backward",k=2)
+
+prediction <- predict(fit,newx = train.sample)
+
+error_table %<>% bind_rows(compute_error_measures(train.sample$corr,prediction,"Backward Sel. AIC","in-sample"))
+
+prediction <- predict(fit,newx = test.sample)
+
+error_table %<>% bind_rows(compute_error_measures(test.sample$corr,prediction,"Backward Sel. AIC","out-of-sample"))
+
+##### Glmnet #####
+set.seed(500)
+glmnet.errors <- lapply(c(0,0.5,1.0),function(alpha){
+
+  # Get output vector for glmnet
+  y <- train.sample %>% pull(all.vars(full.model.formula)[1])
+
+  # Get predictors for glmnet
+  x <- model.matrix(full.model.formula,train.sample)[,-1]
+
+  set.seed(500)
+  fit <- cv.glmnet(x,y,family="gaussian",alpha=alpha)
+
+  prediction <- predict(fit,newx = x,s = "lambda.min")
+
+  train.error <- compute_error_measures(train.sample$corr,prediction,paste0("glmnet, alpha=",alpha),"in-sample")
+
+  xnew <- model.matrix(full.model.formula,test.sample)[,-1]
+
+  prediction <- predict(fit,newx = xnew,s = "lambda.min")
+
+  test.error <- compute_error_measures(test.sample$corr,prediction,paste0("glmnet, alpha=",alpha),"out-of-sample")
+
+  bind_rows(train.error,test.error)
+
+}) %>% bind_rows()
+
+error_table %<>% bind_rows(glmnet.errors)
+
+##### PCR #####
+
+set.seed(100)
+fit <- pcr(full.model.formula,data=train.sample,scale=TRUE,validation="CV", jackknife=TRUE)
+
+ncomps <- c(20,30,selectNcomp(fit,method="onesigma",plot=TRUE)) %>% sort
+
+pcr_errors <- lapply(ncomps,function(ncomp){
+
+  prediction <- predict(fit,newdata = train.sample, type = "response",ncomp = ncomp)
+
+  train.error <- compute_error_measures(train.sample$corr,prediction,paste0("PCR, ncomp=",ncomp),"in-sample")
+
+  prediction <- predict(fit,newx = test.sample,ncomp = ncomp)
+
+  test.error <- compute_error_measures(test.sample$corr,prediction,paste0("PCR, ncomp=",ncomp),"out-of-sample")
+
+  bind_rows(train.error,test.error)
+
+}) %>% bind_rows()
+
+error_table %<>% bind_rows(pcr_errors)
+
+##### PLS #####
+
+set.seed(100)
+fit <- plsr(full.model.formula,data=train.sample,scale=TRUE,validation="CV", jackknife=TRUE)
+summary(fit)
+validationplot(fit)
+
+# 32 components are enough to account for at least 90% of variaiblity in both X and Y. Also use optimal number of components
+
+ncomps <- c(32,selectNcomp(fit,method="onesigma",plot=TRUE)) %>% sort
+
+pls_errors <- lapply(ncomps,function(ncomp){
+
+  prediction <- predict(fit,newdata = train.sample, type = "response",ncomp = ncomp)
+
+  train.error <- compute_error_measures(train.sample$corr,prediction,paste0("PLS, ncomp=",ncomp),"in-sample")
+
+  prediction <- predict(fit,newx = test.sample,ncomp = ncomp)
+
+  test.error <- compute_error_measures(test.sample$corr,prediction,paste0("PLS, ncomp=",ncomp),"out-of-sample")
+
+  bind_rows(train.error,test.error)
+
+}) %>% bind_rows()
+
+error_table %<>% bind_rows(pls_errors)
+
+###### Table 4 #####
+
+# Prediction errors from models to correct bias in indirect estimates of U5M
+
+# Save error table
+
+ft <- error_table  %>% 
+  mutate_at(vars(one_of("rmse","rmedse")),funs(sprintf("%0.6f",.))) %>%
+  flextable() %>% set_header_labels(rmse="Root mean\nsquare error",
+                                    rmedse="Root median\nsquare error",
+                                    mre="Mean relative\nerror",
+                                    medre="Median relative\nerror") %>%
+  merge_v(j=1)
+
+doc <- read_docx() %>%
+  body_add_par(value = "Model selection.", style = "table title") %>% 
+  body_add_flextable(ft)
+
+doc %>% print("tables/table4.docx")
+
+# glmnet with alpha = 1 was chosen to be the best-fitting-model
+
+##### Best fitting model #####
+
+# Get output vector for glmnet
+
+y <- to.model %>% pull(all.vars(full.model.formula)[1])
+
+# Get predictors for glmnet
+
+x <- model.matrix(full.model.formula,to.model)[,-1]
+
+alpha <- 1.0
+
+set.seed(500)
+best.fitting.model <- cv.glmnet(x,y,family="gaussian",alpha=alpha)
+
+lambda <-best.fitting.model$lambda.min
+
+# To use for predictions later.
+
+best.fitting.model <- glmnet(x,y,family="gaussian",alpha=alpha,lambda = lambda)
+
+save(x,best.fitting.model,full.model.formula,file="results/best_fitting_model.RData")
+
+```
+
+## Replication of Figures
+
+```{r, ori-fig-4, fig.cap="Original figure 4"}
+knitr::include_graphics("figures/figure4-ori.png")
+```
+
+```{r, ori-fig-4, fig.cap="Original figure 4"}
+knitr::include_graphics("figures/figure5-ori.png")
+```
+
+The figure 4 (\@ref(fig:ori-fig-4)) and 5 (\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model (REF) to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population.
+By comparing differences in U5M estimates of unadjusted data and adjusted data by two models, the authors suggested that the new model was able to correct the bias from indirect estimates.
+
+With the new simulation data and best-fitting model, we re-generated both figures by using R scripts provided (\@ref(fig:rep-fig-4) and \@ref(fig:rep-fig-5)).
+
+To prepare data for plotting, we created three functions with R code provided by the authors to estimate U5M by indirect method, Ward-Zaba model and our best-performing model respectively.
+These three functions were then further integrated into a function called plot_fig to create our replicated figures.
+
+```{r, indirect-estimates}
+
+indirect_estimate <- function(dataset, fig_country){
+
+  # extract data of the plotting country
+  cntry <- dataset %>% filter(country==fig_country)
+
+  # indirect estimates with empirical data
+  source("R/indirect_estimates_functions_for_reg.R")
+  ind_surv <- ind_est_computations(cntry,2010) %>% as.data.frame() %>% mutate(agegroup=1:7)
+  ind_surv %<>% 
+    select(agegroup,fiveq0=fiveq0, t_ref,refdate)
+
+  # merge indirect estimates based on empirical data from Malawi or Tanzania with
+  # data on HIV, ART and fertility from that same country
+  # ind_est contains the results of indirect_estimates_functions applied to the data from Malawi or Tanzania
+  # cntry contains the data from Malawi or Tanzania (from facevalidity.csv)
+  iep <- cntry %>% 
+    left_join(ind_surv,by="agegroup") %>% 
+    rename(fiveq0_surv=fiveq0) %>% 
+    mutate(agegroup=factor(agegroup)) %>% 
+    select(-cd)
+  colnames(iep) <- c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009","fiveq0_surv", "t_ref","refdate")
+
+  return(iep) 
+}
+
+```
+
+```{r, Ward-Zaba-model}
+wz_model <- function(iep){
+  # Ward & Zaba (2008) coefficients
+  a=c(.1134,.1202,.1107,.1720,.1749,.1504,.2552)
+  b=c(-.0226,-.0438,-.0882,-.1412,-.1758,-.1849,-.3269)
+  c=c(-.1112,-.0978,-.0373,.0163,.042,.1807,.5594)
+
+  # W&Z adjustment factor from applying coefficients to empirical data
+  nz = a*iep$hiv2000+b*(iep$hiv2000^2)+c*iep$prev15to19_2010
+
+  #Convert each n q 0 + n(z) into 5q0
+
+  # Standard indirect estimates, adding Ward&Zaba adj factors
+  # n = c(1,2,3,5,10,15,20)
+  # UN General Female LT, e(0)=60, for above n
+  l = c(.92688,.91025,.90151,.89193,.89534,.89002,.88195)
+  q = 1-l
+  logitMLT=.5*log(q/(1-q))
+  alpha = NA
+  fiveq0WZ = NA
+  Y5 = NA
+
+  nq0 <- ind_est_computations_core(cntry,2010)$nq0
+
+  for(i in 1:7){
+    Y5[i] = .5*log((nq0[i]+nz[i])/(1-(nq0[i]+nz[i])))
+    alpha[i] = Y5[i] - logitMLT[i]
+    fiveq0WZ[i] = exp(2*(alpha[i]+logitMLT[4]))/(1+exp(2*(alpha[i]+logitMLT[4])))
+  }
+
+  # fiveq0WZ are the Ward & Zaba estimates of U5M
+  return(fiveq0WZ)
+}
+```
+
+```{r, best-performing-model}
+best_model <- function(iep, fig_name, best.fitting.model, full.model.formula){
+
+  # Get the new predictions in matrix format
+  nx <- model.matrix(full.model.formula,iep %>% mutate(corr=0))[,-1]
+
+  # Make predictions
+  prediction <- predict(best.fitting.model,newx = nx,se=TRUE,s="lambda.min") %>% 
+    as.data.frame %>% 
+    set_colnames("PredictedAdj")
+
+  # Bootstrap confidence intervals for predictions
+  if(is.null(boot_outs[[fig_name]])){
+    # Get output vector for glmnet
+    y <- to.model %>% pull(all.vars(full.model.formula)[1])
+
+    # Get predictors for glmnet
+    x <- model.matrix(full.model.formula,to.model)[,-1]
+    Xy <-  cbind(y,x)
+    boot_models <- list(length=5)
+    t0 <- predict(best.fitting.model,newx = x,s="lambda.min") %>% as.vector
+    tt <- boot_models[1:5] %>%purrr::map(~predict(best.fitting.model,newx = x,s="lambda.min") %>% as.vector) %>% unlist()
+    #lambda <- best.fitting.model$lambda.min
+    lambda <- best.fitting.model$lambda
+    theta <- as.matrix(coef(best.fitting.model, s = "lambda.min"))
+    pi_hat <- predict(best.fitting.model, newx = x, s = "lambda.min", type = "response")
+    n <- length(pi_hat)
+    y_star <- sapply(seq_len(length(tt)), function(i) ifelse(runif(n) <= pi_hat, 1, 0))
+    suff_stat <- t(y_star) %*% x
+
+    #bcapar(t0=t0,tt=tt,bb=suff_stat)
+    # This is the function that we will use for each bootstrap sample
+    theta <- function(xy,nx,glmnet_alpha,lambda){
+      X <- xy[,2:(dim(xy)[2])]
+      Y <- xy[,1]
+
+      # Fit the model for the sample and using the same lambda as in the best model fitted
+      m <- glmnet(X,Y,family="gaussian",alpha=glmnet_alpha,lambda=lambda)
+
+      # Keep the models computed for each bootstrap sample in order to compute prediction intervals faster
+      # boot_models <<- c(boot_models,list(m))
+
+      # Prediction
+      predict(m,newx = nx,s="lambda.min") %>% as.vector
+
+    }
+
+    # Keep the models computed for each bootstrap sample in order to compute prediction intervals faster
+    # boot_models <- list()
+
+    # Compute bootstrapped confedence intervals for each prediction point
+    boot_outs[[fig_name]] <- lapply(1:nrow(nx),function(i){
+      # i <- 1
+      new_data <- nx[i,,drop=FALSE]
+
+      set.seed(600)
+
+      bcajack(x=Xy,B=500,func=theta,new_data,1.0,lambda)
+    })
+  }
+
+  # Get two tailed 95% CI values 
+  ci <- boot_outs[[fig_name]] %>% lapply(function(r){
+
+    r$lims %>% as.data.frame() %>% mutate(quantile=as.numeric(rownames(.))) %>% filter(quantile %in% c(0.025,0.975)) %>% select(quantile,bca) %>% mutate(quantile=case_when(quantile==0.025 ~ "ci.lo",TRUE~ "ci.hi")) %>% spread(quantile,bca)
+  }) %>% bind_rows()
+
+  prediction %<>% bind_cols(ci)
+  return(prediction)
+}
+```
+
+```{r, figure-plot}
+plot_fig <- function(fig_name, fig_country){
+  # Load empirical data}
+  fv <- fread("./data/facevalidity.csv") %>% set_colnames(c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009")) %>% mutate(cd=ceb-cs)
+
+  # Load bootstrap estimates for model predictions
+  if(file.exists("results/boot_outs.RData")){
+    load("results/boot_outs.RData")
+  }else{
+    boot_outs <- list()  
+  }
+
+  # Load saved best fit model
+  # Contains best.fitting.model and full.model.formula
+  load("results/best_fitting_model.RData")
+
+  # indirect estimates with empirical data
+  iep <- indirect_estimate(fv, fig_country)
+
+  # estimate from Ward & Zaba (2008) model
+  fiveq0WZ <- wz_model(iep)
+
+  # make prediction with best-performing model
+  prediction <- best_model(iep, fig_name, best.fitting.model, full.model.formula)
+
+  # Prepare data for plot
+  to.plot <- bind_cols(iep, prediction,data.frame(fiveq0WZ=fiveq0WZ)) %>% 
+    mutate(fit=fiveq0_surv+PredictedAdj,lwr=fiveq0_surv+ci.lo,upr=fiveq0_surv+ci.hi)%>% # Compute adjusted values and prediction interval upper and lower limits
+    select(refdate,upr,lwr,Adjusted=fit,Unadjusted=fiveq0_surv,`Ward & Zaba`=fiveq0WZ) %>%   # Transform data for plot
+    gather(type,value,-refdate,-upr,-lwr)
+
+  # Plot
+  ggplot(to.plot,aes(x=refdate,y=value,shape=type)) +
+    geom_point(color="black", size=3) +
+    geom_errorbar(aes(ymin=lwr, ymax=upr))+ 
+    xlab("Year") +
+    ylab("Under-five mortality (deaths per live birth)") +
+    scale_shape_discrete(name="Estimate type", # Legend label, use darker colors
+                         labels=c("Adjusted", "Unadjusted","Ward & Zaba"))+
+    labs(title= fig_country) + 
+    theme_classic() +
+    theme(axis.text = element_text(color="black"))
+
+}
+```
+
+```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data"}
+
+# Figure 4
+# Under-five mortality estimates using CEB/CS from 2010 DHS for Malawi:
+# unadjusted (crude), adjusted (corrected) using best-performing model, and using Ward &
+# Zaba (2008) model
+
+plot_fig('figure4','Malawi')
+
+```
+
+```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data"}
+
+# Figure 5
+# Under-five mortality estimates using CEB/CS from 2010 DHS for Tanzania:
+# unadjusted (crude), adjusted (corrected) using best-performing model, and using Ward &
+# Zaba (2008) model
+
+plot_fig('figure5','Tanzania')
+
+```
+
+# Result
+
+# Conclusion
+
+# References cited

### 2020-02-23 analysis/paper.Rmd
add notice for pre-cached data
edit functions for figures

@@
 
-```{r, model-fitting-and-selection, include=FALSE}
+```{r, model-fitting-and-selection, echo=FALSE}
 
@@
 These three functions were then further integrated into a function called plot_fig to create our replicated figures.
+We also included a file with necessary data for calculating the confidence interval by bootstrapping for prediction with our best-fitting model, since this step usually takes more than 40 minutes to finish.
 
@@
 ```{r, best-performing-model}
-best_model <- function(iep, fig_name, best.fitting.model, full.model.formula){
+best_model <- function(iep, fig_name){
+
+  # Load saved best fit model
+  # Contains best.fitting.model and full.model.formula
+  load("results/best_fitting_model.RData")
 
@@
 plot_fig <- function(fig_name, fig_country){
-  # Load empirical data}
+  # Load empirical data
   fv <- fread("./data/facevalidity.csv") %>% set_colnames(c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009")) %>% mutate(cd=ceb-cs)
@@
 
-  # Load saved best fit model
-  # Contains best.fitting.model and full.model.formula
-  load("results/best_fitting_model.RData")
-
   # indirect estimates with empirical data
@@
   # make prediction with best-performing model
-  prediction <- best_model(iep, fig_name, best.fitting.model, full.model.formula)
+  prediction <- best_model(iep, fig_name)
 

### 2020-02-23 analysis/paper.Rmd
edit method and reference of the paper

--- 2020-02-23 analysis/paper.Rmd
@@
 The model was written using the R language [team2013r] and some of the main packages used were tidyverse (v1.2.1) and foreach (v1.4.4).
-The original implementation utilized the doReids package (v1.1.1) that launches a back-end server in Docker for parallel processing that was not adopted in our replication.
-The main limitation is the computing resource we had for this project within a short period of time.
-Based on experiments from multiple team members, finishing the first simulation step would require an 8-core CPU machine for 2 weeks.
-In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we made a decision to reduce the sample size of our model -- the number of populations used in the simulation input has been reduced from 4480 to 1000.
-The code provided by the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented.
+The original implementation utilized the doReids package (v1.1.1) that offers a back-end server for parallel processing that was not adopted in our replication due to the limitation in resources.
+In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we reduced the number of populations used in the simulation input from 4480 to 1000.
+The code provided the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented.
 However, we discovered some mistakes regarding output inconsistency and wrong variable names.
@@
 ## Model Description
+The goal of this large-scale simulation used in the original paper is to build a wide range of demographic histories that accounts for the variability between actual populations, to measure how biases such as the independence assumption of mother and children survival rate in a population influenced by HIV/AIDS would exist.
+@quattrochi2019 used stochastic simulation methods to generate a set of population with input parameters such as demographic features, mortality rate for U5M, maternal mortality rate, HIV prevalence, etc.
+The input simulation parameters vary over a large interval for different age groups across different populations.
+In each time step of one year, each woman in the population has some probabilities to give birth for those who are 15 to 49 years old, being infected with HIV, initiating ART (if HIV-positive), and dying given the simulation parameters.
+Through the aging process in the simulation population, children born with infected HIV mothers will have a probability of getting infected while all children will have a probability of mortality.
+
+## Simulation Construction 
+### Range of Inputs
+
+Since the primary goal in @quattrochi2019 was to measure bias in indirect estimates across a set of populations that have experienced different rates of fertility, mortality, HIV infection, and ART (antiretroviral therapy) initiation.
+We adopted the same method to generate a set of population using the following then inputs: fertility, adult mortality, U5M, percent of 15–19 year olds who are sexually active, maternal mortality in 1990, percent annual decline in the maternal mortality rate, HIV incidence, duration of breastfeeding, and ART coverage.
+The original paper simulated one population for each combination of inputs resulting in a total of 4480 populations.
+Since the original paper did not mention the amount of time required to complete the simulation, we started our process with the same 4480 populations but without parallel computing, we could not complete the simulation within the timeframe of our study and stopped the simulation when it reached 1000.
+For the purpose of replication, we set the seed to be a different value before running the simulation.
+
+### Simulation Parameters
+
+Initially, we attempted to change the year parameter and the population age parameter in the simulation construction process.
+We tried to change the year interval of our study from [1906, 2010] to [1906, 1966] as well as changing the population age interval from [0,120] to [0,75].
+This turned out to be unsuccessful since the the above parameters were not only used in the simulation process but also in the following model selection and regression application.
+For instance, to estimate HIV prevalence, the model divides the population into 7 age groups, which cannot be achieved if the year or age parameter was reduced.
+As a result, we used the same size and date for the initial population in the simulation. We initiated with 22,500 women who were aged 15 years in 1906, and ran the simulation through 2010.
+
+For other parameters in simulation, we followed the original paper and used the same data sources.
+Annual probability of birth was defined as a function of calendar year and mother’s age and set to zero for women younger than 15 years and older than 49 years.
+Estimates of age-specific fertility rates (ASFR) from the United Nations Population Division’s World Fertility Data [@unitednations] was used and for years when ASFR were not available, the adjusted the nearest available ASFR using the interpolated estimates of the total fertility rate (TFR) from the United Nations Population Division’s World Population Prospects [@unitednations_2013] was used.
+
+As mentioned in the original paper, according to [@chen2010fertility], women between 15–19 years old who were HIV-positive experienced higher ASFRs compared to HIV-negative women, depending if they were sexually active.
+Therefore the ASFR parameter was adjusted using the ratios estimated by [@chen2010fertility].
+ASFR was also adjusted based on the studies that have found an increase in pregnancy following ART [@myer2010impact][@makumbi2011associations][@homsy2009reproductive].
+The effect of ART on fertility likely depends on a variety of factors including age, cluster of differentiation 4 (CD4) count at initiation according to the original paper.
+For the simulation model, the assumption that the ASFR for 15–19 year olds is not affected by ART was made to further simplify the simulation.
+Inputs relating to maternal mortality was calculated using MMR (maternal deaths per 100,000 births) in 1990 and the annual decline in MMR since 1990.
+The initial value of the MMR according to [@hogan2010maternal] was either 0.0012 or 0.012 and the annual rate of decline was set to 0 or 7.3% and calculated as:
+
+For other inputs including the annual probability of HIV infection[@hogan2012spline][@heuveline2003hiv], the parameters governing CD4[@hallett2008estimating], ART coverage[@databank], the annual probability of death[@worldhealthorganization_2013] and the probability of mother-to-child transmission of HIV[@stover2008spectrum], we also used the same data sources and derivation methods.
 

### 2020-02-23 analysis/paper.Rmd
add code for simulation

@@
 knitr::opts_chunk$set(echo = TRUE)
+source("R/simulation_functions.R")
+source("R/indirect_estimates_functions_revised.R")
 ```
@@
 
-```{r, load-packages}
+```{r, load-packages, include=FALSE}
+install.packages('http://cran.r-project.org/src/contrib/Archive/tidyverse/tidyverse_1.2.1.tar.gz',repos=NULL, type="source")
+packageVersion("tidyverse")
+require(tidyverse)
+# ── Attaching packages ───────────────────────────────────────────────────────────────────────────── tidyverse 1.2.1 ──
+# ✔ ggplot2 3.1.0       ✔ purrr   0.3.0  
+# ✔ tibble  2.0.1       ✔ dplyr   0.8.0.1
+# ✔ tidyr   0.8.2       ✔ stringr 1.4.0  
+# ✔ readr   1.1.1       ✔ forcats 0.3.0  
+# ── Conflicts ──────────────────────────────────────────────────────────────────────────────── tidyverse_conflicts() ──
+# ✖ dplyr::filter() masks stats::filter()
+# ✖ dplyr::lag()    masks stats::lag()
 
+install.packages('https://cran.r-project.org/src/contrib/Archive/foreach/foreach_1.4.4.tar.gz',repos=NULL, type="source")
+packageVersion("foreach")
+# 1.4.4
+require(foreach)
+
+install.packages('https://cran.r-project.org/src/contrib/Archive/iterators/iterators_1.0.10.tar.gz',repos=NULL, type="source")
+packageVersion("iterators")
+# 1.0.10
+require(iterators)
 ```
@@
 ## Simulation Construction 
+```{r eval = FALSE}
+##### DEFINE FOLDERS #######
+
+if(!dir.exists("results/figdata")){
+  dir.create("results/figdata",recursive = TRUE)
+}
+
+if(!dir.exists("results/regdata")){
+  dir.create("results/regdata",recursive = TRUE)
+}
+
+# Simulation results will be stored here
+
+if(!dir.exists("results/models")){
+  dir.create("results/models",recursive = TRUE)
+}
+
+##### SIMULATION #####
+
+# Read data
+
+hivhogan = read.csv("data/inc_curves.csv",head=TRUE)
+mort_series = read.csv("data/IHME_female_mortSMALL.csv",head=TRUE)
+adultmort = read.csv("data/MLTfemSMALL.csv",head=TRUE)
+worldfert= read.csv("data/world_fert.csv",head=TRUE)
+tfr_series= read.csv("data/tfr_gapminder_long.csv",head=TRUE)
+art_series= read.csv("data/sampleART.csv",head=TRUE)
+u5m_edit= read.csv("data/u5m_edit.csv",head=TRUE)
+matmort = read.csv("data/matmort.csv",head=TRUE)
+
+# Add HIV-free populations to hivhogan
+hivhogan$Country <- factor(hivhogan$Country,levels=c(levels(hivhogan$Country),"zero"))
+hivhogan[63,1] = "zero"
+hivhogan[63,c(2:47)]=0
+
+# Create parameters sets
+
+fertcountry <- c("Botswana","Uganda")
+cm_cntry <-  c("Mali","Morocco")
+am_cntry <-  c("Madagascar","Sudan")
+sexactive15 <-  c(30,70)
+mmr0 <- c(0.0012,0.012)
+mmr_dec <- c(0,0.073) # annual percent decline in MMR
+curve <- c("BotswanaUrban","LesothoRural","MalawiRural","UgandaRural","CamerounRural","zero","BotUrb2x")
+bfeed <- c(6,18)
+art_col <- c("zero","Botswana","Cameroon","Malawi","Bot_dub")
+growth <-  c(0.03)
+yrend <- 2010
+# Starting year
+yrstart <- 1906
+
+# CD4 threshold
+
+threshold <- 200
+
+# expand.grid() will create dataset with unique combinations in each row
+
+inputs <- expand.grid(fertcountry, cm_cntry, am_cntry, sexactive15,mmr0,mmr_dec,curve,bfeed,art_col,growth,yrend,yrstart,threshold)
+names(inputs) = c("fertcountry", "cm_cntry","am_cntry","sexactive15","mmr0","mmr_dec","curve","bfeed","art_col","growth","yrend","yrstart","threshold")
+
+# Save the inputs set for later
+
+save(inputs,file = "results/models/inputs.RData")
+
+file_number_format <- paste0("%0",nchar(as.character(nrow(inputs))),"d")
+
+# Set size of initial population
+
+initial_populations<- c(7500,15000,22500)
+#initial_populations<- c(100,150,225)
+years <- c(1906:2010)
+#years <- c(1906:1966)
+ages <- c(0:120)
+#ages <- c(0:75)
+
+# Run the simulations for each initial population
+
+for(ip in initial_populations){
+
+  # Parameter sets already processed
+  #ip = 150
+  psets_already <- as.integer(gsub("models|\\.Rdata","",list.files(paste0("./results/models/p",ip),pattern = "models\\d+",full.names = FALSE)))
+
+  psets_to_process <- setdiff(1:nrow(inputs),psets_already)
+
+  # Run simulation for each set of parameters
+
+  foreach(pset=psets_to_process,.packages = c("dplyr")) %dopar%{
+
+    inp <- inputs[pset,]
+
+    source("R/simulation_functions.R")
+
+    source("R/indirect_estimates_functions.R")
+
+    # Run simmulation
+    set.seed(1)
+    results <- bigsim(inp,initialpop=ip,years,ages,hivhogan,mort_series,adultmort,worldfert,tfr_series,art_series,u5m_edit,matmort)
+
+    # Save simulation results
+    #save(meta.res1,file=file.path(paste0("./results/models"),paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))
+
+    save(results,file=file.path(paste0("./results/models/p",ip),paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))
+  }
+
+}
+
+# Collect hiv2000 differences among initial populations
+
+# For each input row
+
+hiv2000 <- foreach(pset=1:1000,.packages = c("dplyr")) %dopar%{
+
+  # And each initial population
+  paste0("p",initial_populations) %>% lapply(function(popini){
+
+    # Load the results file for that initial population and input set
+
+    load(file.path("results/models/",popini,paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))
+    print(paste(initial_populations,pset))
+    # Extract the value of hiv2000
+    r <- results[[1]]$hiv2000
+
+    names(r) <- popini
+    r
+  }) %>% unlist() %>% t %>% data.frame(input_set=pset,.) # Put everything in one data.frame
+
+} %>% bind_rows
+
+# Compute difference
+
+hiv2000 <- hiv2000 %>% mutate(p7500_minus_p15000=p7500-p15000,p15000_minus_p22500=p15000-p22500)
+
+save(hiv2000,file = "results/hiv2000.RData")
+
+##### INDIRECT ESTIMATES FOR INITAL POPULATION 22500 #####
+
+# See if we have something not processed
+
+psets_already <- intersect(as.integer(gsub("figdata.|\\.Rdata","",list.files("./results/figdata/p22500",pattern = "figdata.\\d+",full.names = FALSE))),
+                           as.integer(gsub("regdata.|\\.Rdata","",list.files("./results/regdata/p22500",pattern = "regdata.\\d+",full.names = FALSE))))
+
+psets_to_process <- setdiff(1:nrow(inputs),psets_already)
+
+# For each input set not processed
+
+foreach(pset=1:1000,.packages = c("dplyr")) %dopar%{
+
+  source("R/indirect_estimates_functions_revised.R")
+
+  load(file.path("results/models/p22500/",paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))
+  print(pset)
+
+  w <- results[[7]]
+
+  # Are babies of HIV moms getting HIV?
+  # What percent of HIV moms babies have HIV?
+  # How many born with HIV? 
+  # df = as.data.frame(w)
+  # y <- subset(df, hiv_date==2009)
+  # plot(df$dob, df$hiv_date)  
+  ##################################################
+
+  # format data
+  momkidsclean <- as.data.frame(w)
+
+  # Using all surviving women
+  isf <-ind_est(momkidsclean)
+
+  # Using all women
+  isf_all <-ind_est_all(momkidsclean)
+
+  #Using surviving women and women who died from HIV
+  isf_hiv <- ind_est_hiv(momkidsclean)
+
+  ie.three <- cbind(isf,isf_all,isf_hiv,pset)
+
+  ###################################################################
+  # Save results for regressions and for figures
+  ###################################################################
+
+  inp.plus <- as.data.frame(results[1])    
+  regdata <- cbind(ie.three,inp.plus[1,], row.names = NULL) # Data for regressions
+  figdata <- results[c(2:6,8:9)] # Data for figures
+
+  save(regdata,file=file.path("results/regdata/p22500/",paste("regdata.",sprintf(file_number_format,pset),".Rdata",sep="")))
+  save(figdata,file=file.path("results/figdata/p22500/",paste("figdata.",sprintf(file_number_format,pset),".Rdata",sep="")))
+  print('Saved')
+  rm(results,regdata,figdata,momkidsclean,isf,isf_all,isf_hiv,ie.three)
+}
+
+##### COLLECT DATA FROM ALL SIMULATIONS FOR INITIAL POPULATION 22500 #####
+
+# Merge regdata
+
+psets_to_process <- as.integer(gsub("regdata.|\\.Rdata","",list.files("./results/regdata/p22500",pattern = "regdata.\\d+",full.names = FALSE))) %>% sort
+
+nbd2k <- psets_to_process %>% lapply(function(pset){
+
+  filename <- file.path("results/regdata/p22500",paste("regdata.",sprintf(file_number_format,pset),".Rdata",sep=""))
+  tryCatch({
+    load(filename)
+
+    regdata %>% mutate(i=pset) %>% select(i,everything())
+
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+}) %>% bind_rows() %>% as.data.frame()
+
+save(nbd2k,file="results/regdata/p22500/regdata_all.Rdata")
+
+# Merge figdata
+
+h <- list()
+tf<- list()
+ar<- list()
+hd<- list()
+u5m <- list()
+inps <- list()
+arp <- list()
+hivinc <- list()
+
+psets_to_process <- as.integer(gsub("figdata.|\\.Rdata","",list.files("results/figdata/p22500",pattern = "figdata.\\d+",full.names = FALSE))) %>% sort
+
+for(pset in psets_to_process){
+
+  filename <- file.path("results/figdata/p22500",paste("figdata.",sprintf(file_number_format,pset),".Rdata",sep=""))
+  tryCatch({
+    load(filename)
+
+    h[[pset]] <- figdata[[3]]
+    tf[[pset]] <- figdata[[1]]
+    ar[[pset]] <- figdata[[2]]
+    hd[[pset]] <- figdata[[4]]
+    inps[[pset]] <- figdata[[5]]
+    arp[[pset]] <- figdata[[6]]
+    hivinc[[pset]] <- figdata[[7]]
+    rm(figdata)
+
+  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
+}
+
+save(h,tf,ar,hd,u5m,inps,arp,file="results/figdata/p22500/figdata_all.Rdata")
+```
 ### Range of Inputs
@@
 save(x,best.fitting.model,full.model.formula,file="results/best_fitting_model.RData")
-
 ```

### 2020-02-23 analysis/paper.Rmd
Revise overview section with updated simulation time estimation

--- 2020-02-23 analysis/paper.Rmd
@@
 The model was written using the R language [team2013r] and some of the main packages used were tidyverse (v1.2.1) and foreach (v1.4.4).
-The original implementation utilized the doReids package (v1.1.1) that offers a back-end server for parallel processing that was not adopted in our replication due to the limitation in resources.
-In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we reduced the number of populations used in the simulation input from 4480 to 1000.
-The code provided the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented.
+The original implementation utilized the doReids package (v1.1.1) that launches a back-end server in Docker for parallel processing that was not adopted in our replication.
+The main limitation is the computing resource we had for this project within a short period of time.
+Based on experiments performed on a m4.xlarge EC2 instance on AWS, the original simulation settings in the target paper consumed over 63% CPU and 4GB of memory and kept running for over 3 days using Docker doRedis package.
+The operator estimated of finishing the entire simulation step would require an 8-core CPU machine more than 1 week.
+In order to complete the simulation for constructing the model in a reasonable time frame without parallel computing, we had to stop the simulation and reduced the number of populations used in the simulation input from 4480 to 1000.
+The code provided by the author on Github (https://github.com/jquattro/hiv-childmort-bias) was mostly well-organized and documented.
 However, we discovered some mistakes regarding output inconsistency and wrong variable names.
@@
 ## Model Description
-The goal of this large-scale simulation used in the original paper is to build a wide range of demographic histories that accounts for the variability between actual populations, to measure how biases such as the independence assumption of mother and children survival rate in a population influenced by HIV/AIDS would exist.
-@quattrochi2019 used stochastic simulation methods to generate a set of population with input parameters such as demographic features, mortality rate for U5M, maternal mortality rate, HIV prevalence, etc.
-The input simulation parameters vary over a large interval for different age groups across different populations.
-In each time step of one year, each woman in the population has some probabilities to give birth for those who are 15 to 49 years old, being infected with HIV, initiating ART (if HIV-positive), and dying given the simulation parameters.
-Through the aging process in the simulation population, children born with infected HIV mothers will have a probability of getting infected while all children will have a probability of mortality.
-
-## Simulation Construction 
-### Range of Inputs
-
-Since the primary goal in @quattrochi2019 was to measure bias in indirect estimates across a set of populations that have experienced different rates of fertility, mortality, HIV infection, and ART (antiretroviral therapy) initiation.
-We adopted the same method to generate a set of population using the following then inputs: fertility, adult mortality, U5M, percent of 15–19 year olds who are sexually active, maternal mortality in 1990, percent annual decline in the maternal mortality rate, HIV incidence, duration of breastfeeding, and ART coverage.
-The original paper simulated one population for each combination of inputs resulting in a total of 4480 populations.
-Since the original paper did not mention the amount of time required to complete the simulation, we started our process with the same 4480 populations but without parallel computing, we could not complete the simulation within the timeframe of our study and stopped the simulation when it reached 1000.
-For the purpose of replication, we set the seed to be a different value before running the simulation.
-
-### Simulation Parameters
-
-Initially, we attempted to change the year parameter and the population age parameter in the simulation construction process.
-We tried to change the year interval of our study from [1906, 2010] to [1906, 1966] as well as changing the population age interval from [0,120] to [0,75].
-This turned out to be unsuccessful since the the above parameters were not only used in the simulation process but also in the following model selection and regression application.
-For instance, to estimate HIV prevalence, the model divides the population into 7 age groups, which cannot be achieved if the year or age parameter was reduced.
-As a result, we used the same size and date for the initial population in the simulation. We initiated with 22,500 women who were aged 15 years in 1906, and ran the simulation through 2010.
-
-For other parameters in simulation, we followed the original paper and used the same data sources.
-Annual probability of birth was defined as a function of calendar year and mother’s age and set to zero for women younger than 15 years and older than 49 years.
-Estimates of age-specific fertility rates (ASFR) from the United Nations Population Division’s World Fertility Data [@unitednations] was used and for years when ASFR were not available, the adjusted the nearest available ASFR using the interpolated estimates of the total fertility rate (TFR) from the United Nations Population Division’s World Population Prospects [@unitednations_2013] was used.
-
-As mentioned in the original paper, according to [@chen2010fertility], women between 15–19 years old who were HIV-positive experienced higher ASFRs compared to HIV-negative women, depending if they were sexually active.
-Therefore the ASFR parameter was adjusted using the ratios estimated by [@chen2010fertility].
-ASFR was also adjusted based on the studies that have found an increase in pregnancy following ART [@myer2010impact][@makumbi2011associations][@homsy2009reproductive].
-The effect of ART on fertility likely depends on a variety of factors including age, cluster of differentiation 4 (CD4) count at initiation according to the original paper.
-For the simulation model, the assumption that the ASFR for 15–19 year olds is not affected by ART was made to further simplify the simulation.
-Inputs relating to maternal mortality was calculated using MMR (maternal deaths per 100,000 births) in 1990 and the annual decline in MMR since 1990.
-The initial value of the MMR according to [@hogan2010maternal] was either 0.0012 or 0.012 and the annual rate of decline was set to 0 or 7.3% and calculated as:
-
-For other inputs including the annual probability of HIV infection[@hogan2012spline][@heuveline2003hiv], the parameters governing CD4[@hallett2008estimating], ART coverage[@databank], the annual probability of death[@worldhealthorganization_2013] and the probability of mother-to-child transmission of HIV[@stover2008spectrum], we also used the same data sources and derivation methods.
 

### 2020-02-23 analysis/paper.Rmd
update

--- 2020-02-23 analysis/paper.Rmd
@@
 Through the aging process in the simulation population, children born with infected HIV mothers will have a probability of getting infected while all children will have a probability of mortality.
+For the calculation of bias, we used the formula provided in the original paper. 
+$$ Relative bias = {\frac{IE_{survivors} - IE_{survivors&HIV deaths}}{IE_{survivors&HIV deaths}} $$        
+$$ Absolute bias = IE_{survivors} - IE_{survivors&HIV deaths}$$    
+According to [@quattrochi2019], ${IE_{survivors}$ was the estimates of U5M based on living women in 2010, and IE_{survivors&HIV deaths} was the estimates of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
 

### 2020-02-23 analysis/paper.Rmd
update
edit based on comments

--- 2020-02-23 analysis/paper.Rmd
@@
 initial_populations<- c(7500,15000,22500)
-#initial_populations<- c(100,150,225)
 years <- c(1906:2010)
-#years <- c(1906:1966)
 ages <- c(0:120)
-#ages <- c(0:75)
 
@@
     # Save simulation results
-    #save(meta.res1,file=file.path(paste0("./results/models"),paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))
-
     save(results,file=file.path(paste0("./results/models/p",ip),paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))

### 2020-02-24 analysis/paper.Rmd

@@
 Through the aging process in the simulation population, children born with infected HIV mothers will have a probability of getting infected while all children will have a probability of mortality.
+
+For the calculation of bias, we used the formula provided in the original paper. 
+$$ Relative bias = {\frac{IE_{survivors} - IE_{survivors&HIV deaths}}{IE_{survivors&HIV deaths}} $$        
+$$ Absolute bias = IE_{survivors} - IE_{survivors&HIV deaths}$$    
+According to [@quattrochi2019], ${IE_{survivors}$ was the estimates of U5M based on living women in 2010, and IE_{survivors&HIV deaths} was the estimates of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
 

### 2020-02-24 analysis/paper.Rmd
debug making figures

--- 2020-02-23 analysis/paper.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-source("R/simulation_functions.R")
-source("R/indirect_estimates_functions_revised.R")
+knitr::opts_knit$set(root.dir = '../') 
 ```
@@
 ```{r, load-packages, include=FALSE}
+if(!require(tidyverse)){
   install.packages('http://cran.r-project.org/src/contrib/Archive/tidyverse/tidyverse_1.2.1.tar.gz',repos=NULL, type="source")
+}
 packageVersion("tidyverse")
@@
 
+if(!require(foreach)){
   install.packages('https://cran.r-project.org/src/contrib/Archive/foreach/foreach_1.4.4.tar.gz',repos=NULL, type="source")
+}
 packageVersion("foreach")
@@
 
+if(!require(iterators)){
   install.packages('https://cran.r-project.org/src/contrib/Archive/iterators/iterators_1.0.10.tar.gz',repos=NULL, type="source")
+}
 packageVersion("iterators")
@@
 require(iterators)
+
+if(!require(bcaboot)){
+  install.packages("bcaboot",dependencies = TRUE,repos='http://cran.us.r-project.org')
+}
+require(bcaboot)
+packageVersion("bcaboot")
+# 0.2.1
+
+if(!require(data.table)){
+  install.packages("data.table",dependencies = TRUE,repos='http://cran.us.r-project.org')
+}
+require(data.table)
+packageVersion("data.table")
+# 1.11.8
+
+if(!require(pls)){
+  install.packages("pls",dependencies = TRUE,repos='http://cran.us.r-project.org')
+}
+require(pls)
+packageVersion("pls")
+# 2.7.0
+
+if(!require(glmnet)){
+  install.packages('https://cran.r-project.org/src/contrib/Archive/glmnet/glmnet_2.0-16.tar.gz',repos=NULL, type="source")
+}
+require(glmnet)
+packageVersion("glmnet")
+# 2.0.16
+
+if(!require(officer)){
+  install.packages("officer",dependencies = TRUE,repos='http://cran.us.r-project.org')
+}
+require(officer)
+packageVersion("officer")
+# 0.3.2
+
+if(!require(flextable)){
+  install.packages("flextable",dependencies = TRUE,repos='http://cran.us.r-project.org')
+}
+require(flextable)
+packageVersion("flextable")
+# 0.5.1
+
+if(!require(magrittr)){
+  install.packages("magrittr",dependencies = TRUE,repos='http://cran.us.r-project.org')
+}
+require(magrittr)
+packageVersion("magrittr")
+# 1.5
+
 ```
@@
 ## Simulation Construction 
-```{r eval = FALSE}
+```{r, include = FALSE, eval=FALSE}
 ##### DEFINE FOLDERS #######
@@
 
-```{r, model-fitting-and-selection, echo=FALSE}
+```{r, model-fitting-and-selection, include=FALSE}
 
@@
 
-```{r, ori-fig-4, fig.cap="Original figure 4"}
-knitr::include_graphics("figures/figure4-ori.png")
+```{r, ori-fig-4, fig.cap="Original figure 4", echo=FALSE}
+knitr::include_graphics("figure4-ori.png")
 ```
 
-```{r, ori-fig-4, fig.cap="Original figure 4"}
-knitr::include_graphics("figures/figure5-ori.png")
+```{r, ori-fig-5, fig.cap="Original figure 5", echo=FALSE}
+knitr::include_graphics("figure5-ori.png")
 ```
 
-The figure 4 (\@ref(fig:ori-fig-4)) and 5 (\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model (REF) to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population.
+The figure 4 (Fig.
+\@ref(fig:ori-fig-4)) and 5 (Fig.
+\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model (REF) to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population.
 By comparing differences in U5M estimates of unadjusted data and adjusted data by two models, the authors suggested that the new model was able to correct the bias from indirect estimates.
 
-With the new simulation data and best-fitting model, we re-generated both figures by using R scripts provided (\@ref(fig:rep-fig-4) and \@ref(fig:rep-fig-5)).
+With the new simulation data and best-fitting model, we re-generated both figures by using R scripts provided (Fig. \@ref(fig:rep-fig-4) and  Fig. \@ref(fig:rep-fig-5)).
 
@@
 
-```{r, indirect-estimates}
+```{r, indirect-estimates, include=FALSE}
 
@@
 
-```{r, Ward-Zaba-model}
-wz_model <- function(iep){
+```{r, Ward-Zaba-model, include=FALSE}
+wz_model <- function(dataset, fig_country, iep){
+
+  # extract data of the plotting country
+  cntry <- dataset %>% filter(country==fig_country)
+
   # Ward & Zaba (2008) coefficients
@@
 
-```{r, best-performing-model}
+```{r, best-performing-model, include=FALSE}
 best_model <- function(iep, fig_name){
@@
   load("results/best_fitting_model.RData")
+
+  # Load bootstrap estimates for model predictions
+  if(file.exists("results/boot_outs.RData")){
+    load("results/boot_outs.RData")
+  }else{
+    boot_outs <- list()  
+  }
 
@@
 
-```{r, figure-plot}
+```{r, figure-plot, include=FALSE}
 plot_fig <- function(fig_name, fig_country){
@@
   fv <- fread("./data/facevalidity.csv") %>% set_colnames(c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009")) %>% mutate(cd=ceb-cs)
-
-  # Load bootstrap estimates for model predictions
-  if(file.exists("results/boot_outs.RData")){
-    load("results/boot_outs.RData")
-  }else{
-    boot_outs <- list()  
-  }
 
@@
   # estimate from Ward & Zaba (2008) model
-  fiveq0WZ <- wz_model(iep)
+  fiveq0WZ <- wz_model(fv, fig_country, iep)
 
@@
 
-```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data"}
+```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data", echo=FALSE， fig.align='center'}
 
@@
 
-```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data"}
+```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data", echo=FALSE， fig.align='center'}
 

### 2020-02-24 analysis/paper.Rmd
add result section to paper

@@
 
-```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data", echo=FALSE， fig.align='center'}
+```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data", echo=FALSE, fig.align='center'}
 
@@
 
-```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data", echo=FALSE， fig.align='center'}
+```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data", echo=FALSE, fig.align='center'}
 
@@
 
+Our implementation of the model to estimate the bias in U5M in HIV/AIDS affected populations from @quattrochi2019 was able to replicate the process and output target tables and figures.
+Consistently, indirect methods underestimate U5M in older populations, while little difference was shown in younger age groups between unadjusted and adjusted estimates.
+Applying our model to 2010 survey data from Malawi and Tanzania, we show that indirect methods would underestimate U5M by up to 12%.
+
+Table 1 (Fig.
+\@ref(fig:table1)) compares the prediction errors across the 13 models in the same format as in Table 4 in the original paper.
+Compared to the estimation errors calculated in the original paper, we had very similar results of the root mean square errors and root median square errors for all models.
+However, our in-sample mean relative errors were consistently greater compared to the original paper and we suspect this could be due to the reduction in population size during the simulation.
+Consistent with the original paper, no single model in the table showed clear advantages for all the metrics over others we selected the generalized linear regression with alpha equal to 1 as our predictive model since it showed the best performance for most of the metrics.
+
+```{r, table1, fig.cap="Table 1", echo=FALSE}
+knitr::include_graphics("table1_re.png")
+```
+
+Table 2 (Fig.
+\@ref(fig:table2)) shows our estimation by applying the best performing model on 1,000 simulated populations and consistent with the original implementation, the mean HIV prevalence we calculated among women aged 15–49 across populations was 7% in 1990, 15% in 2000, and 11% in 2010.
+The highest HIV prevalence in any simulation was 40% in 2000.
+Our results related to ART showed the same trend as in the original paper that both prevalence and coverage increase with time, although the exact numbers varied.
+The mean ART coverage across our simulations was 0% in 2004 and 8% in 2010.
+The mean ART prevalence was < 0.1% in 2005 and 0.19% in 2009.
+The highest ART prevalence in any simulation was 2.3% in 2009 and this was 1.7% lower compared to the original model.
+The mean TFR across simulations was 4.88 in 2000 and 4.28 in 2010, also consistent with the original model.
+
+```{r, table2, fig.cap="Table 2", echo=FALSE}
+knitr::include_graphics("table2_re.png")
+```
+
+Table 3 (Fig.
+\@ref(fig:table3)) shows the bias in indirect estimates across age groups.
+According to the original paper, negative numbers indicate that unadjusted estimates were lower than adjusted estimates.
+The mean absolute bias was largest for estimates from women aged 35–39 and 40–44 (−0.021) and smallest for estimates from women aged 15–19 (0.000).
+Our results showed the same age group with a small variance in the exact numbers.
+The original paper also emphasized on the largest absolute bias recorded, which was −0.071 for estimates from women 35–39 in our calculation and -0.069 for them.
+The paper pointed out that this means that the estimated U5M was 71 deaths per 1000 live births lower when using only reports from surviving women compared to reports from surviving women and women who died from HIV/AIDS.
+
+```{r, table3, fig.cap="Table 3", echo=FALSE}
+knitr::include_graphics("table3_re.png")
+```
+
+Based on our results, we believe that our effort in replicating @quattrochi2019 was successful despite some inconsistency.
+We argue that the differences could be due to the smaller population size we used in our simulation though further investigation would be needed to discover the exact causes.
+
 # Conclusion

### 2020-02-24 analysis/paper.Rmd
update

--- 2020-02-24 analysis/paper.Rmd
@@
 knitr::opts_chunk$set(echo = TRUE)
-knitr::opts_knit$set(root.dir = '../') 
 ```
@@
 
-```{r, load-packages, include=FALSE}
-if(!require(tidyverse)){
-  install.packages('http://cran.r-project.org/src/contrib/Archive/tidyverse/tidyverse_1.2.1.tar.gz',repos=NULL, type="source")
-}
-packageVersion("tidyverse")
-require(tidyverse)
-# ── Attaching packages ───────────────────────────────────────────────────────────────────────────── tidyverse 1.2.1 ──
-# ✔ ggplot2 3.1.0       ✔ purrr   0.3.0  
-# ✔ tibble  2.0.1       ✔ dplyr   0.8.0.1
-# ✔ tidyr   0.8.2       ✔ stringr 1.4.0  
-# ✔ readr   1.1.1       ✔ forcats 0.3.0  
-# ── Conflicts ──────────────────────────────────────────────────────────────────────────────── tidyverse_conflicts() ──
-# ✖ dplyr::filter() masks stats::filter()
-# ✖ dplyr::lag()    masks stats::lag()
-
-if(!require(foreach)){
-  install.packages('https://cran.r-project.org/src/contrib/Archive/foreach/foreach_1.4.4.tar.gz',repos=NULL, type="source")
-}
-packageVersion("foreach")
-# 1.4.4
-require(foreach)
-
-if(!require(iterators)){
-  install.packages('https://cran.r-project.org/src/contrib/Archive/iterators/iterators_1.0.10.tar.gz',repos=NULL, type="source")
-}
-packageVersion("iterators")
-# 1.0.10
-require(iterators)
-
-if(!require(bcaboot)){
-  install.packages("bcaboot",dependencies = TRUE,repos='http://cran.us.r-project.org')
-}
-require(bcaboot)
-packageVersion("bcaboot")
-# 0.2.1
-
-if(!require(data.table)){
-  install.packages("data.table",dependencies = TRUE,repos='http://cran.us.r-project.org')
-}
-require(data.table)
-packageVersion("data.table")
-# 1.11.8
-
-if(!require(pls)){
-  install.packages("pls",dependencies = TRUE,repos='http://cran.us.r-project.org')
-}
-require(pls)
-packageVersion("pls")
-# 2.7.0
-
-if(!require(glmnet)){
-  install.packages('https://cran.r-project.org/src/contrib/Archive/glmnet/glmnet_2.0-16.tar.gz',repos=NULL, type="source")
-}
-require(glmnet)
-packageVersion("glmnet")
-# 2.0.16
-
-if(!require(officer)){
-  install.packages("officer",dependencies = TRUE,repos='http://cran.us.r-project.org')
-}
-require(officer)
-packageVersion("officer")
-# 0.3.2
-
-if(!require(flextable)){
-  install.packages("flextable",dependencies = TRUE,repos='http://cran.us.r-project.org')
-}
-require(flextable)
-packageVersion("flextable")
-# 0.5.1
-
-if(!require(magrittr)){
-  install.packages("magrittr",dependencies = TRUE,repos='http://cran.us.r-project.org')
-}
-require(magrittr)
-packageVersion("magrittr")
-# 1.5
+```{r, load-packages}
 
@@
 For the calculation of bias, we used the formula provided in the original paper. 
-$$ Relative bias = {\frac{IE_{survivors} - IE_{survivors&HIV deaths}}{IE_{survivors&HIV deaths}} $$        
-$$ Absolute bias = IE_{survivors} - IE_{survivors&HIV deaths}$$    
-According to [@quattrochi2019], ${IE_{survivors}$ was the estimates of U5M based on living women in 2010, and IE_{survivors&HIV deaths} was the estimates of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
+$$Relative bias = \frac{IE_{survivors} - IE_{survivors,HIV deaths}}{IE_{survivors,HIV deaths}}$$
+
+$$Absolute bias = IE_{survivors} - IE_{survivors,HIV deaths}$$
+According to [@quattrochi2019], $IE_{survivor}$ was the estimates of U5M based on living women in 2010, and $IE_{survivors,HIV deaths}$ was the estimates of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
+$a_{d}$
 
 ## Simulation Construction 
-```{r, include = FALSE, eval=FALSE}
-##### DEFINE FOLDERS #######
-
-if(!dir.exists("results/figdata")){
-  dir.create("results/figdata",recursive = TRUE)
-}
-
-if(!dir.exists("results/regdata")){
-  dir.create("results/regdata",recursive = TRUE)
-}
-
-# Simulation results will be stored here
-
-if(!dir.exists("results/models")){
-  dir.create("results/models",recursive = TRUE)
-}
-
-##### SIMULATION #####
-
-# Read data
-
-hivhogan = read.csv("data/inc_curves.csv",head=TRUE)
-mort_series = read.csv("data/IHME_female_mortSMALL.csv",head=TRUE)
-adultmort = read.csv("data/MLTfemSMALL.csv",head=TRUE)
-worldfert= read.csv("data/world_fert.csv",head=TRUE)
-tfr_series= read.csv("data/tfr_gapminder_long.csv",head=TRUE)
-art_series= read.csv("data/sampleART.csv",head=TRUE)
-u5m_edit= read.csv("data/u5m_edit.csv",head=TRUE)
-matmort = read.csv("data/matmort.csv",head=TRUE)
-
-# Add HIV-free populations to hivhogan
-hivhogan$Country <- factor(hivhogan$Country,levels=c(levels(hivhogan$Country),"zero"))
-hivhogan[63,1] = "zero"
-hivhogan[63,c(2:47)]=0
-
-# Create parameters sets
-
-fertcountry <- c("Botswana","Uganda")
-cm_cntry <-  c("Mali","Morocco")
-am_cntry <-  c("Madagascar","Sudan")
-sexactive15 <-  c(30,70)
-mmr0 <- c(0.0012,0.012)
-mmr_dec <- c(0,0.073) # annual percent decline in MMR
-curve <- c("BotswanaUrban","LesothoRural","MalawiRural","UgandaRural","CamerounRural","zero","BotUrb2x")
-bfeed <- c(6,18)
-art_col <- c("zero","Botswana","Cameroon","Malawi","Bot_dub")
-growth <-  c(0.03)
-yrend <- 2010
-# Starting year
-yrstart <- 1906
-
-# CD4 threshold
-
-threshold <- 200
-
-# expand.grid() will create dataset with unique combinations in each row
-
-inputs <- expand.grid(fertcountry, cm_cntry, am_cntry, sexactive15,mmr0,mmr_dec,curve,bfeed,art_col,growth,yrend,yrstart,threshold)
-names(inputs) = c("fertcountry", "cm_cntry","am_cntry","sexactive15","mmr0","mmr_dec","curve","bfeed","art_col","growth","yrend","yrstart","threshold")
-
-# Save the inputs set for later
-
-save(inputs,file = "results/models/inputs.RData")
-
-file_number_format <- paste0("%0",nchar(as.character(nrow(inputs))),"d")
-
-# Set size of initial population
-
-initial_populations<- c(7500,15000,22500)
-years <- c(1906:2010)
-ages <- c(0:120)
-
-# Run the simulations for each initial population
-
-for(ip in initial_populations){
-
-  # Parameter sets already processed
-  #ip = 150
-  psets_already <- as.integer(gsub("models|\\.Rdata","",list.files(paste0("./results/models/p",ip),pattern = "models\\d+",full.names = FALSE)))
-
-  psets_to_process <- setdiff(1:nrow(inputs),psets_already)
-
-  # Run simulation for each set of parameters
-
-  foreach(pset=psets_to_process,.packages = c("dplyr")) %dopar%{
-
-    inp <- inputs[pset,]
-
-    source("R/simulation_functions.R")
-
-    source("R/indirect_estimates_functions.R")
-
-    # Run simmulation
-    set.seed(1)
-    results <- bigsim(inp,initialpop=ip,years,ages,hivhogan,mort_series,adultmort,worldfert,tfr_series,art_series,u5m_edit,matmort)
-
-    # Save simulation results
-    save(results,file=file.path(paste0("./results/models/p",ip),paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))
-  }
-
-}
-
-# Collect hiv2000 differences among initial populations
-
-# For each input row
-
-hiv2000 <- foreach(pset=1:1000,.packages = c("dplyr")) %dopar%{
-
-  # And each initial population
-  paste0("p",initial_populations) %>% lapply(function(popini){
-
-    # Load the results file for that initial population and input set
-
-    load(file.path("results/models/",popini,paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))
-    print(paste(initial_populations,pset))
-    # Extract the value of hiv2000
-    r <- results[[1]]$hiv2000
-
-    names(r) <- popini
-    r
-  }) %>% unlist() %>% t %>% data.frame(input_set=pset,.) # Put everything in one data.frame
-
-} %>% bind_rows
-
-# Compute difference
-
-hiv2000 <- hiv2000 %>% mutate(p7500_minus_p15000=p7500-p15000,p15000_minus_p22500=p15000-p22500)
-
-save(hiv2000,file = "results/hiv2000.RData")
-
-##### INDIRECT ESTIMATES FOR INITAL POPULATION 22500 #####
-
-# See if we have something not processed
-
-psets_already <- intersect(as.integer(gsub("figdata.|\\.Rdata","",list.files("./results/figdata/p22500",pattern = "figdata.\\d+",full.names = FALSE))),
-                           as.integer(gsub("regdata.|\\.Rdata","",list.files("./results/regdata/p22500",pattern = "regdata.\\d+",full.names = FALSE))))
-
-psets_to_process <- setdiff(1:nrow(inputs),psets_already)
-
-# For each input set not processed
-
-foreach(pset=1:1000,.packages = c("dplyr")) %dopar%{
-
-  source("R/indirect_estimates_functions_revised.R")
-
-  load(file.path("results/models/p22500/",paste("models",sprintf(file_number_format,pset),".Rdata",sep="")))
-  print(pset)
-
-  w <- results[[7]]
-
-  # Are babies of HIV moms getting HIV?
-  # What percent of HIV moms babies have HIV?
-  # How many born with HIV? 
-  # df = as.data.frame(w)
-  # y <- subset(df, hiv_date==2009)
-  # plot(df$dob, df$hiv_date)  
-  ##################################################
-
-  # format data
-  momkidsclean <- as.data.frame(w)
-
-  # Using all surviving women
-  isf <-ind_est(momkidsclean)
-
-  # Using all women
-  isf_all <-ind_est_all(momkidsclean)
-
-  #Using surviving women and women who died from HIV
-  isf_hiv <- ind_est_hiv(momkidsclean)
-
-  ie.three <- cbind(isf,isf_all,isf_hiv,pset)
-
-  ###################################################################
-  # Save results for regressions and for figures
-  ###################################################################
-
-  inp.plus <- as.data.frame(results[1])    
-  regdata <- cbind(ie.three,inp.plus[1,], row.names = NULL) # Data for regressions
-  figdata <- results[c(2:6,8:9)] # Data for figures
-
-  save(regdata,file=file.path("results/regdata/p22500/",paste("regdata.",sprintf(file_number_format,pset),".Rdata",sep="")))
-  save(figdata,file=file.path("results/figdata/p22500/",paste("figdata.",sprintf(file_number_format,pset),".Rdata",sep="")))
-  print('Saved')
-  rm(results,regdata,figdata,momkidsclean,isf,isf_all,isf_hiv,ie.three)
-}
-
-##### COLLECT DATA FROM ALL SIMULATIONS FOR INITIAL POPULATION 22500 #####
-
-# Merge regdata
-
-psets_to_process <- as.integer(gsub("regdata.|\\.Rdata","",list.files("./results/regdata/p22500",pattern = "regdata.\\d+",full.names = FALSE))) %>% sort
-
-nbd2k <- psets_to_process %>% lapply(function(pset){
-
-  filename <- file.path("results/regdata/p22500",paste("regdata.",sprintf(file_number_format,pset),".Rdata",sep=""))
-  tryCatch({
-    load(filename)
-
-    regdata %>% mutate(i=pset) %>% select(i,everything())
-
-  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
-}) %>% bind_rows() %>% as.data.frame()
-
-save(nbd2k,file="results/regdata/p22500/regdata_all.Rdata")
-
-# Merge figdata
-
-h <- list()
-tf<- list()
-ar<- list()
-hd<- list()
-u5m <- list()
-inps <- list()
-arp <- list()
-hivinc <- list()
-
-psets_to_process <- as.integer(gsub("figdata.|\\.Rdata","",list.files("results/figdata/p22500",pattern = "figdata.\\d+",full.names = FALSE))) %>% sort
-
-for(pset in psets_to_process){
-
-  filename <- file.path("results/figdata/p22500",paste("figdata.",sprintf(file_number_format,pset),".Rdata",sep=""))
-  tryCatch({
-    load(filename)
-
-    h[[pset]] <- figdata[[3]]
-    tf[[pset]] <- figdata[[1]]
-    ar[[pset]] <- figdata[[2]]
-    hd[[pset]] <- figdata[[4]]
-    inps[[pset]] <- figdata[[5]]
-    arp[[pset]] <- figdata[[6]]
-    hivinc[[pset]] <- figdata[[7]]
-    rm(figdata)
-
-  }, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})
-}
-
-save(h,tf,ar,hd,u5m,inps,arp,file="results/figdata/p22500/figdata_all.Rdata")
-```
 ### Range of Inputs
@@
 save(x,best.fitting.model,full.model.formula,file="results/best_fitting_model.RData")
+
 ```
@@
 
-```{r, ori-fig-4, fig.cap="Original figure 4", echo=FALSE}
-knitr::include_graphics("figure4-ori.png")
+```{r, ori-fig-4, fig.cap="Original figure 4"}
+knitr::include_graphics("figures/figure4-ori.png")
 ```
 
-```{r, ori-fig-5, fig.cap="Original figure 5", echo=FALSE}
-knitr::include_graphics("figure5-ori.png")
+```{r, ori-fig-4, fig.cap="Original figure 4"}
+knitr::include_graphics("figures/figure5-ori.png")
 ```
 
-The figure 4 (Fig.
-\@ref(fig:ori-fig-4)) and 5 (Fig.
-\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model (REF) to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population.
+The figure 4 (\@ref(fig:ori-fig-4)) and 5 (\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model (REF) to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population.
 By comparing differences in U5M estimates of unadjusted data and adjusted data by two models, the authors suggested that the new model was able to correct the bias from indirect estimates.
 
-With the new simulation data and best-fitting model, we re-generated both figures by using R scripts provided (Fig. \@ref(fig:rep-fig-4) and  Fig. \@ref(fig:rep-fig-5)).
+With the new simulation data and best-fitting model, we re-generated both figures by using R scripts provided (\@ref(fig:rep-fig-4) and \@ref(fig:rep-fig-5)).
 
@@
 
-```{r, indirect-estimates, include=FALSE}
+```{r, indirect-estimates}
 
@@
 
-```{r, Ward-Zaba-model, include=FALSE}
-wz_model <- function(dataset, fig_country, iep){
-
-  # extract data of the plotting country
-  cntry <- dataset %>% filter(country==fig_country)
-
+```{r, Ward-Zaba-model}
+wz_model <- function(iep){
   # Ward & Zaba (2008) coefficients
@@
 
-```{r, best-performing-model, include=FALSE}
+```{r, best-performing-model}
 best_model <- function(iep, fig_name){
@@
   load("results/best_fitting_model.RData")
-
-  # Load bootstrap estimates for model predictions
-  if(file.exists("results/boot_outs.RData")){
-    load("results/boot_outs.RData")
-  }else{
-    boot_outs <- list()  
-  }
 
@@
 
-```{r, figure-plot, include=FALSE}
+```{r, figure-plot}
 plot_fig <- function(fig_name, fig_country){
@@
   fv <- fread("./data/facevalidity.csv") %>% set_colnames(c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009")) %>% mutate(cd=ceb-cs)
+
+  # Load bootstrap estimates for model predictions
+  if(file.exists("results/boot_outs.RData")){
+    load("results/boot_outs.RData")
+  }else{
+    boot_outs <- list()  
+  }
 
@@
   # estimate from Ward & Zaba (2008) model
-  fiveq0WZ <- wz_model(fv, fig_country, iep)
+  fiveq0WZ <- wz_model(iep)
 
@@
 
-```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data", echo=FALSE， fig.align='center'}
+```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data"}
 
@@
 
-```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data", echo=FALSE， fig.align='center'}
+```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data"}
 

### 2020-02-24 analysis/paper.Rmd

--- 2020-02-24 analysis/paper.Rmd
@@
 For the calculation of bias, we used the formula provided in the original paper. 
-$$ Relative bias = {\frac{IE_{survivors} - IE_{survivors&HIV deaths}}{IE_{survivors&HIV deaths}} $$        
-$$ Absolute bias = IE_{survivors} - IE_{survivors&HIV deaths}$$    
-According to [@quattrochi2019], ${IE_{survivors}$ was the estimates of U5M based on living women in 2010, and IE_{survivors&HIV deaths} was the estimates of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
+$$Relative bias = \frac{IE_{survivors} - IE_{survivors,HIV deaths}}{IE_{survivors,HIV deaths}}$$
+
+$$Absolute bias = IE_{survivors} - IE_{survivors,HIV deaths}$$
+According to [@quattrochi2019], $IE_{survivor}$ was the estimates of U5M based on living women in 2010, and $IE_{survivors,HIV deaths}$ was the estimates of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
+$a_{d}$
 

### 2020-02-24 analysis/paper.Rmd
minor change

@@
 For the calculation of bias, we used the formula provided in the original paper. 
-$$Relative bias = \frac{IE_{survivors} - IE_{survivors,HIV deaths}}{IE_{survivors,HIV deaths}}$$
+$$\mbox{Relative bias} = \frac{IE_{survivors} - IE_{survivors,HIV deaths}}{IE_{survivors,HIV deaths}}$$
 
-$$Absolute bias = IE_{survivors} - IE_{survivors,HIV deaths}$$
+$$\mbox{Absolute bias} = IE_{survivors} - IE_{survivors,HIV deaths}$$
 According to [@quattrochi2019], $IE_{survivor}$ was the estimates of U5M based on living women in 2010, and $IE_{survivors,HIV deaths}$ was the estimates of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
@@
 Consistently, indirect methods underestimate U5M in older populations, while little difference was shown in younger age groups between unadjusted and adjusted estimates.
-Applying our model to 2010 survey data from Malawi and Tanzania, we show that indirect methods would underestimate U5M by up to 12%.
+Applying our model to 2010 survey data from Malawi and Tanzania, we show that indirect methods would underestimate U5M by up to 7%.
 

### 2020-02-24 analysis/paper.Rmd
add debug code in the regression and add some explanation

@@
 
+We revised some parts of the regression model and found inconsistent variable names during the debugging process for the reproduction of the tables and figures to be sucessful.
+In the following section of the original regression code, we changed the parameter in the bootstrapping model from 1:500 to 1:5 to make it run faster.
+We also caught the inconsistency in the variable  “lambda.min” where the parameter should be “lambda” from the model.
+
+```{r, eval=FALSE}
+ boot_models <- list(length=5)
+    t0 <- predict(best.fitting.model,newx = x,s="lambda.min") %>% as.vector
+    tt <- boot_models[1:5] %>%purrr::map(~predict(best.fitting.model,newx = x,s="lambda.min") %>% as.vector) %>% unlist()
+    #lambda <- best.fitting.model$lambda.min
+    lambda <- best.fitting.model$lambda
+    theta <- as.matrix(coef(best.fitting.model, s = "lambda.min"))
+    pi_hat <- predict(best.fitting.model, newx = x, s = "lambda.min", type = "response")
+    n <- length(pi_hat)
+    y_star <- sapply(seq_len(length(tt)), function(i) ifelse(runif(n) <= pi_hat, 1, 0))
+    suff_stat <- t(y_star) %*% x
+```
 # Result

### 2020-02-24 analysis/paper.Rmd

@@
 ```{r, include = FALSE, eval=FALSE}
+# This section of code is designed to be no executable as it takes a long time to generate simulations.
+# Instead, input data was provided as .RData files
+
 ##### DEFINE FOLDERS #######
@@
 # Conclusion
+Our goal of this project is to replicate a published research project and allow readers to produce identical figures to those in the target paper.
+We were able to accomplish this goal by replicating the key figures in the original paper using R and publishing well-structured source code along with data source in a licensed Github repository.
+This paper.md R Markdown file will be used as the main entry point of our project and is fully executable.
+However, we intentionally set a couple code chunks with evaluation as False since simulating those data sources does not fit within the timeframe for any group who is trying to reproduce them.
+Instead, we created .RData file as a battery-included solution for future reproducers.
+
+We leveraged some useful Reproducibility techniques introduced in the class such as Tidyverse libraries and R Markdown syntax.
+The in-file executable R chunks come extremely handy during our reproducible process.
+
+There are, however, challenges throughout our explorations.
+The first major issue we encountered was the lengthy source data generation step which required extensive research and deep dive into source code.
+Given the nature of the HIV/AIDS estimation research, the data samples are required to be large enough and comprehensive which should cover across multiple geological locations and a full range of populations.
+Therefore, the original authors designed the raw data to be generated on-the-go rather than archived in open source.
+Our team underestimated the challenge of reproducing all sources by the time we picked this article.
+To tackle this, all team members gathered together and agreed to make tradeoffs - reducing population size during simulation.
+The decision was made in favor of the balance between quality of our research and a timely deliverable.
+We identified this as an area to improve (focus on verification of data source before replication) during our retrospective of this project.
+Another challenge was using R in an unfamiliar field of research.
+Given none of us had previous domain knowledge regarding HIV/AIDS research, we spent a lot of time understanding the terminologies while exploring the rationale behind the R code used in the original paper.
+We had multiple issues trying to reproduce a specific failure in R from multiple environments.
+For example, we still couldn’t fully understand some of the issues during “knit” failures which had been resolved by cleaning up cache in R Studio.
+We strongly believe that using Docker or a shared development environment could dramatically improve our developer experience where all bugs could be reproduced easily no matter where the Docker image is running.
+
+In general, we gained hands-on experience from reproducing an entire research project.
+After rounds of reviews and merging conflicts, we developed our own Github usage convention and code styles.
+We would like to carry these invaluable skills to our future Data Science journey.
 

### 2020-02-24 analysis/paper.Rmd
add one reference

@@
 
-## Simulation Construction
-
 ## Model Fitting and Model Selection
@@
 \@ref(fig:ori-fig-4)) and 5 (Fig.
-\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model (REF) to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population.
+\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model [@ward2008effect] to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population.
 By comparing differences in U5M estimates of unadjusted data and adjusted data by two models, the authors suggested that the new model was able to correct the bias from indirect estimates.
@@
 
-```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data", echo=FALSE, fig.align='center'}
+```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data", echo=FALSE}
 
@@
 
-```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data", echo=FALSE, fig.align='center'}
+```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data", echo=FALSE}
 

### 2020-02-24 analysis/paper.Rmd
change fig size

@@
 
-```{r, ori-fig-4, fig.cap="Original figure 4", echo=FALSE}
+```{r, ori-fig-4, fig.cap="Original figure 4", echo=FALSE, out.width="10%"}
 knitr::include_graphics("figure4-ori.png")
@@
 
-```{r, ori-fig-5, fig.cap="Original figure 5", echo=FALSE}
+```{r, ori-fig-5, fig.cap="Original figure 5", echo=FALSE, out.width="10%"}
 knitr::include_graphics("figure5-ori.png")
@@
 
-```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data", echo=FALSE}
+```{r, rep-fig-4, fig.cap="Replication of figure 4 with new simulation data", echo=FALSE, , fig.width=7, fig.height=4}
 
@@
 
-```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data", echo=FALSE}
+```{r, rep-fig-5, fig.cap="Replication of figure 5 with new simulation data", echo=FALSE, , fig.width=7, fig.height=4}
 

### 2020-02-24 analysis/paper.Rmd
fix typos and add missing refs

--- 2020-02-24 analysis/paper.Rmd
@@
 ## Model Description
-The goal of this large-scale simulation used in the original paper is to build a wide range of demographic histories that accounts for the variability between actual populations, to measure how biases such as the independence assumption of mother and children survival rate in a population influenced by HIV/AIDS would exist.
-@quattrochi2019 used stochastic simulation methods to generate a set of population with input parameters such as demographic features, mortality rate for U5M, maternal mortality rate, HIV prevalence, etc.
+The goal of the large-scale simulation used in the original paper was to build a demographic history that accounted for the variability among actual populations to measure if biases such as the independence assumption of mother and children survival rate in a population influenced by HIV/AIDS existed.
+@quattrochi2019 used stochastic simulation methods to generate a set of populations with input parameters including demographic features, mortality rate for U5M, maternal mortality rate, HIV prevalence, etc.
 The input simulation parameters vary over a large interval for different age groups across different populations.
-In each time step of one year, each woman in the population has some probabilities to give birth for those who are 15 to 49 years old, being infected with HIV, initiating ART (if HIV-positive), and dying given the simulation parameters.
+In each time step of one year, each woman in a population has some probabilities to give birth (for women between 15 to 49 years old), being infected with HIV, initiating ART (if HIV-positive), and dying given the simulation parameters.
 Through the aging process in the simulation population, children born with infected HIV mothers will have a probability of getting infected while all children will have a probability of mortality.
 
-For the calculation of bias, we used the formula provided in the original paper. 
+For calculating relative and absolute bias, we used the formulas provided in the original paper. 
 $$\mbox{Relative bias} = \frac{IE_{survivors} - IE_{survivors,HIV deaths}}{IE_{survivors,HIV deaths}}$$
@@
 $$\mbox{Absolute bias} = IE_{survivors} - IE_{survivors,HIV deaths}$$
-According to [@quattrochi2019], $IE_{survivor}$ was the estimates of U5M based on living women in 2010, and $IE_{survivors,HIV deaths}$ was the estimates of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
+According to [@quattrochi2019], $IE_{survivor}$ was the estimattion of U5M based on living women in 2010, and $IE_{survivors,HIV deaths}$ was the estimation of U5M based on living women in 2010 and women died from HIV/AIDS before 2010 given the condition that they would have been 15-49 in 2010 if they had survived.
 $a_{d}$
@@
 Since the primary goal in @quattrochi2019 was to measure bias in indirect estimates across a set of populations that have experienced different rates of fertility, mortality, HIV infection, and ART (antiretroviral therapy) initiation.
-We adopted the same method to generate a set of population using the following then inputs: fertility, adult mortality, U5M, percent of 15–19 year olds who are sexually active, maternal mortality in 1990, percent annual decline in the maternal mortality rate, HIV incidence, duration of breastfeeding, and ART coverage.
+We adopted the same method to generate a set of populations using the following inputs: fertility, adult mortality, U5M, percent of 15–19 year olds who are sexually active, maternal mortality in 1990, percent annual decline in the maternal mortality rate, HIV incidence, duration of breastfeeding, and ART coverage.
 The original paper simulated one population for each combination of inputs resulting in a total of 4480 populations.
-Since the original paper did not mention the amount of time required to complete the simulation, we started our process with the same 4480 populations but without parallel computing, we could not complete the simulation within the timeframe of our study and stopped the simulation when it reached 1000.
-For the purpose of replication, we set the seed to be a different value before running the simulation.
+Since the original paper did not provide information on the amount of time required to complete the simulation, we started our process with the same 4480 populations but without parallel computing, we could not complete the simulation within the timeframe of our study and stopped the simulation when it reached 1000 populations.
+For the purpose of our replication study, we set the seed for simulation to be 1, which was different from the seed value 1000 in the original paper.
 
@@
 
-Initially, we attempted to change the year parameter and the population age parameter in the simulation construction process.
-We tried to change the year interval of our study from [1906, 2010] to [1906, 1966] as well as changing the population age interval from [0,120] to [0,75].
+Initially, we attempted to some parameters in the simulation construction process.
+We tried to reduce the year interval of our study from [1906, 2010] to [1906, 1966] as well as the population age interval from [0,120] to [0,75].
 This turned out to be unsuccessful since the the above parameters were not only used in the simulation process but also in the following model selection and regression application.
 For instance, to estimate HIV prevalence, the model divides the population into 7 age groups, which cannot be achieved if the year or age parameter was reduced.
-As a result, we used the same size and date for the initial population in the simulation. We initiated with 22,500 women who were aged 15 years in 1906, and ran the simulation through 2010.
+As a result, we used the same size and date for the initial population in the simulation.
+We initiated with 22,500 women who were aged 15 years in 1906, and ran the simulation through 2010.
 
@@
 Inputs relating to maternal mortality was calculated using MMR (maternal deaths per 100,000 births) in 1990 and the annual decline in MMR since 1990.
-The initial value of the MMR according to [@hogan2010maternal] was either 0.0012 or 0.012 and the annual rate of decline was set to 0 or 7.3% and calculated as:
+The initial value of the MMR according to [@hogan2010maternal] was either 0.0012 or 0.012 and the annual rate of decline was set to 0 or 7.3%.
 
@@
 
-Newly-generated simulation data was then imported and fitted to various models to select for a best-performing model which provides least errors.
-
-The simulated data was first divided into training set (80%) and test set (20%).
-The training set was fitted to eight different models, which were further evaluated by calculating several error measures (root mean square error, root median square error, mean relative error and median relative error) on both training set (in-sample) and test set (out-of-sample) shown in Table 4.
-
-The glmnet with alpha = 1 was selected as the best-performing model based on its error measures and was used in predictions later.
+Newly-generated simulation data were then imported and fitted to various models to select for a best-performing model and the selection process used least errors to represent model performances.
+
+The simulated data were first divided into training set (80%) and test set (20%).
+The training set was fitted using eight different models, which were further evaluated by calculating several error metrics (root mean square error, root median square error, mean relative error and median relative error) on both training set (in-sample) and test set (out-of-sample) shown in Table 1(Fig.
+\@ref(fig:table1)).
+
+Compared to the estimation errors calculated in the original paper, we had very similar results of the root mean square errors and root median square errors for all models.
+However, our in-sample mean relative errors were consistently greater compared to the original paper and we suspect this could be due to the reduction in population size during the simulation.
+Consistent with the original paper, no single model in the table showed clear advantages for all the metrics over others we selected the generalized linear regression with alpha equal to 1 as our predictive model since it showed the best performance for most of the metrics.
+
+```{r, table1, fig.cap="Table 1", echo=FALSE}
+knitr::include_graphics("table1_re.png")
+```
 
@@
 
-The figure 4 (Fig.
+Figure 4 (Fig.
 \@ref(fig:ori-fig-4)) and 5 (Fig.
-\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model [@ward2008effect] to empirical data from 2010 DHS for Malawi and Tanzania respectively to estimate U5M of different age groups of population.
+\@ref(fig:ori-fig-5)) in the original article were created by applying the best-performing model as well as an independent model [@ward2008effect] to the empirical data from 2010 DHS for Malawi and Tanzania [@nationalbureauofstatistics-nbs_tanzania_icf_macro_2011][@nationalstatisticaloffice_icf_macro_2011] respectively to estimate U5M of different age groups of population.
 By comparing differences in U5M estimates of unadjusted data and adjusted data by two models, the authors suggested that the new model was able to correct the bias from indirect estimates.
@@
 
-To prepare data for plotting, we created three functions with R code provided by the authors to estimate U5M by indirect method, Ward-Zaba model and our best-performing model respectively.
+To prepare data for plotting, we created three functions with R code provided by the authors to estimate U5M using the indirect method, the Ward-Zaba model and our best-performing model.
 These three functions were then further integrated into a function called plot_fig to create our replicated figures.
@@
 
-We revised some parts of the regression model and found inconsistent variable names during the debugging process for the reproduction of the tables and figures to be sucessful.
-In the following section of the original regression code, we changed the parameter in the bootstrapping model from 1:500 to 1:5 to make it run faster.
-We also caught the inconsistency in the variable  “lambda.min” where the parameter should be “lambda” from the model.
+We revised some parts of the regression model and corrected inconsistent variable names in the original code during the debugging process.
+For instance, caught corrected the variable name “lambda.min” where it should have been “lambda”.
 
@@
 
-Table 1 (Fig.
-\@ref(fig:table1)) compares the prediction errors across the 13 models in the same format as in Table 4 in the original paper.
-Compared to the estimation errors calculated in the original paper, we had very similar results of the root mean square errors and root median square errors for all models.
-However, our in-sample mean relative errors were consistently greater compared to the original paper and we suspect this could be due to the reduction in population size during the simulation.
-Consistent with the original paper, no single model in the table showed clear advantages for all the metrics over others we selected the generalized linear regression with alpha equal to 1 as our predictive model since it showed the best performance for most of the metrics.
-
-```{r, table1, fig.cap="Table 1", echo=FALSE}
-knitr::include_graphics("table1_re.png")
-```
-
 Table 2 (Fig.
@@
 # Conclusion
-Our goal of this project is to replicate a published research project and allow readers to produce identical figures to those in the target paper.
+Our goal of this project was to replicate a published research paper and reproduce figures in it.
 We were able to accomplish this goal by replicating the key figures in the original paper using R and publishing well-structured source code along with data source in a licensed Github repository.
-This paper.md R Markdown file will be used as the main entry point of our project and is fully executable.
-However, we intentionally set a couple code chunks with evaluation as False since simulating those data sources does not fit within the timeframe for any group who is trying to reproduce them.
+This paper.Rmd R Markdown file is fully executable therefore can be used as the main entry point to our project.
+However, we intentionally set a couple code chunks with evaluation as False since simulating those data sources does not fit within the timeframe for any group trying to reproduce them.
 Instead, we created .RData file as a battery-included solution for future reproducers.
 
-We leveraged some useful Reproducibility techniques introduced in the class such as Tidyverse libraries and R Markdown syntax.
-The in-file executable R chunks come extremely handy during our reproducible process.
+We leveraged some useful reproducibility techniques introduced in the class such as Tidyverse libraries and R Markdown syntax.
+The in-file executable R chunks come extremely handy during our reproduction process.
 
 There are, however, challenges throughout our explorations.
-The first major issue we encountered was the lengthy source data generation step which required extensive research and deep dive into source code.
-Given the nature of the HIV/AIDS estimation research, the data samples are required to be large enough and comprehensive which should cover across multiple geological locations and a full range of populations.
+The first major issue we encountered was the lengthy source data generation step which required extensive research and deep dive into the source code.
+Given the nature of the HIV/AIDS estimation research, the data samples required had to be comprehensive enough to cover multiple geological locations and a wide range of populations.
 Therefore, the original authors designed the raw data to be generated on-the-go rather than archived in open source.
-Our team underestimated the challenge of reproducing all sources by the time we picked this article.
-To tackle this, all team members gathered together and agreed to make tradeoffs - reducing population size during simulation.
-The decision was made in favor of the balance between quality of our research and a timely deliverable.
-We identified this as an area to improve (focus on verification of data source before replication) during our retrospective of this project.
+Our team initially underestimated the challenge of reproducing all sources and this was also due to the lack of information on completion time in the original paper.
+We had to reduce population size during simulation to ensure the quality of our research and a timely deliverable.
+Looking back, we can identify this as an area that we could improve (focusing on verification of data source before replication).
 Another challenge was using R in an unfamiliar field of research.
-Given none of us had previous domain knowledge regarding HIV/AIDS research, we spent a lot of time understanding the terminologies while exploring the rationale behind the R code used in the original paper.
-We had multiple issues trying to reproduce a specific failure in R from multiple environments.
-For example, we still couldn’t fully understand some of the issues during “knit” failures which had been resolved by cleaning up cache in R Studio.
-We strongly believe that using Docker or a shared development environment could dramatically improve our developer experience where all bugs could be reproduced easily no matter where the Docker image is running.
-
-In general, we gained hands-on experience from reproducing an entire research project.
+Given none of us had previous domain knowledge regarding HIV/AIDS research, we spent a lot of time to understand the terminologies while exploring the rationales behind the R code used in the original paper.
+We had issues trying to understnad specific failures in R givern that multiple team members had different R environment.
+For example, we still couldn’t fully understand some of the “knit” failures which had been resolved by cleaning up cache in R Studio.
+We strongly believe that using Docker or a shared development environment could drastically improve our developer experience where most errors could be reproduced easily no matter where the Docker image is running.
+
+In general, we gained hands-on experience from reproducing a research project.
 After rounds of reviews and merging conflicts, we developed our own Github usage convention and code styles.

### 2020-02-25 analysis/paper.Rmd
change data format

--- 2020-02-24 analysis/paper.Rmd
@@
 author: "Sihao Miao, Ruian Yang, Xiaolu Qian, Jiyu Wang, David Wei"
-date: "`r format(Sys.time(), '%d %B, %Y')`"
+date: "24 February, 2020"
 Repository:

### 2020-03-13 analysis/paper.Rmd
re-organize files and edit rmd file

@@
 # Get data from simulation
-load("./results/regdata_all.Rdata")
-load("./results/inputs.RData")
+load("results/regdata_all.Rdata")
+load("results/inputs.RData")
 
@@
   # Load empirical data
-  fv <- fread("./data/facevalidity.csv") %>% set_colnames(c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009")) %>% mutate(cd=ceb-cs)
+  fv <- fread("data/facevalidity.csv") %>% set_colnames(c("country","agegroup", "ceb","cs","tfr2010","tfr2000","hiv1990", "hiv2000", "hiv2010","art2005","art2006","art2007","art2008","art2009","art2010","art2011","art_prev2005","art_prev2006","art_prev2007","art_prev2008","art_prev2009","prop15to19_2010","prev15to19_2010","hiv2005","hiv2006","hiv2007","hiv2008","hiv2009")) %>% mutate(cd=ceb-cs)
 
